---
title: "DA_communitycomposition"
author: "Lieke"
date: "2024-06-20"
output: html_document
---

```{r packages}
# general
library(tidyverse)
# dissimilarity matrix, NMDS, PERMANOVA, PERMDISP
library(vegan)
# models
library(glmmTMB)
library(mgcv) 
# indicator species
library(indicspecies)
# visualisation
library(patchwork)
library(ggforce) # for NMDS amoebae
library(flextable)
library(gridExtra) # for grid.arrange
library(report)
library(fantaxtic)
library(cowplot)
library(ggnested)
landusetypes <- c("Food_forest", "Forest", "Grassland", "Arable_field")
```

```{r functions}
source("functions.R")
```

```{r data}
dissimlist <- readRDS("dissimlist.rds")

codes_info <- read.csv("Data/codes_info.csv")
sites_info <- codes_info %>%
  select(site, location, landuse) %>%
  distinct()

# species abundance matrices (community data table)
specablist <- readRDS("specablist.rds")

# read in data used for stacked bars taxonomy
taxonomic <- readRDS("taxonomic.rds")

# original data
harvestmen_raw <- read.csv("Data/Harvestmen_P2.csv")
dflist <- readRDS("dflist.rds")

# taxonomic info
taxonomy <- read.csv("Data/macrofauna_specieslist.csv")

# # read in NMDS results
# NMDSreslist <- readRDS("NMDSreslist.rds")
```

# NMDS

```{r NMDS}
# list to save plots in
NMDS_plots <- list()

# list to save results in
NMDSreslist <- list() # will be chill to have this saved instead of having to rerun nmds all the time

for(m in names(dissimlist)){
  df <- dissimlist[[m]]
  set.seed(2983847)
  NMDS_res <- metaMDS(df, k = 2, trymax = 999, trace = FALSE) 
  # autotransform by default TRUE
  
  p <- myNMDS(df, NMDS_res, codes_info) + 
    # temporary for figure
    labs(x = "NMDS1", y = "NMDS2")
  
  NMDS_plots[[m]] <- p
  
  NMDSreslist[[m]] <- NMDS_res

}

saveRDS(NMDSreslist, "NMDSreslist.rds")
# (NMDSplot <- do.call("grid.arrange", c(NMDS_plots, ncol = 3)))
(NMDSplot <- do.call("plot_grid", c(NMDS_plots[c("df_trap", "df_sitesum")], ncol=2, align="hv")))
ggsave("NMDSplot.png", NMDSplot, width = 12, height = 8)
ggsave("NMDSplot.svg", NMDSplot, width = 16, height = 8)


## if we only include communities with at least 5 individuals and only include land use types for which at least 6 observations remain, we can still make the comparison with forests, but not with grass- and croplands (check if at site level this would be remedied)

# NMDS_plots_sep <- list()
# 
# for(m in names(dissimlist)){
#   set.seed(2983847)
#   df <- as.matrix(dissimlist[[m]])
# 
#   rn <- data.frame(code = rownames(df)) %>%
#     left_join(codes_info %>% select(code, landuse, location))
#   
#   # count number of observations per land use type
#   n_obs <- rn %>%
#     group_by(landuse) %>%
#     summarise(n = n()) %>%
#     # now already remove land use type for which there are less than 6 observations
#     filter(n >= 6)
#   
#   # filter out rownames (and column names) of land use type with too few observations
#   rn <- rn %>%
#     filter(landuse %in% n_obs$landuse)
#   
#   df <- df[rn$code, rn$code]
# 
#     
#   p <- myNMDS(df, codes_info) + 
#     # temporary for figure
#     labs(x = "NMDS1", y = "NMDS2")
#   
#   NMDS_plots_sep[[m]] <- p
# 
# }
# 
# 
# 
# (NMDSplot_sep <- do.call("grid.arrange", c(NMDS_plots_sep, ncol = 3)))
# ggsave("NMDSplot_sep.svg", NMDSplot_sep, width = 16, height = 16)
```

# Indicator species

```{r indicspecies}

for(m in names(specablist)){
  set.seed(2983847)
  df <- (specablist[[m]])
  
  site_class <- data.frame(code = rownames(df)) %>%
    left_join(codes_info) %>%
    select(landuse)

indval <- multipatt(df, site_class$landuse, 
                    control = how(nperm=999)) 

#summary(indval) # for all three cases only Phalangium opilio associated with combination of grass- and cropland land use types
summary(indval, alpha = 0.1) # alpha = 1 to display all
}


```

# Taxonomy

Stacked bars with proportion of individuals belonging to certain taxonomic group.
This is in any case at the site level, not trap level, and taking the mean proportion per site or the proportion after having summed two traps per site gives exactly the same results (logically, but also tested it to be sure).

```{r taxonomy_bars_layout}

taxon_harvestmen <- readRDS("taxon_macroarthropods.rds") %>%
  filter(group == "harvestmen")


# colours
pal <- c("#0C7BDC", "#FFC20A","#40B0A6","#5D3A9B","#FEFE62","#DC3220","#1AFF1A","#D35FB7","#994F00","#88CCEE")

# theme
set_theme <- list(
  theme_bw(),
  theme(panel.grid = element_blank(),
        strip.background.x = element_blank(), 
        strip.placement = "outside",
        axis.line = element_line(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.length.x = unit(0, "pt"),
        panel.border = element_blank(),
        panel.spacing.x = unit(7, "pt"), # does what it implies, space the panels of the facet 
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        strip.text.x = element_text(size = 14))#,
        # legend.text = element_text(size = 13), 
        # legend.title = element_text(size = 14))
)

```

```{r taxonomy_bars}
# mean per site
tax_stack_mean <- taxon_harvestmen %>%
  left_join(sites_info) %>%
  mutate(landuse = factor(landuse, levels = c("Food_forest", "Forest",
                                              "Grassland", "Cropland"))) %>%
  filter(mean_quant > 0) %>%
  ggnested(aes(x = site, y = mean_quant, main_group = family, sub_group = genus), 
         main_palette = pal) + 
  geom_col(position = position_fill(reverse = TRUE)) + 
  # 0 at level 0 of y axis
  scale_y_continuous(expand = expansion(mult = c(0, .05))) + #, labels = scales::percent
  # add padding around outer bars
  scale_x_discrete(expand = expansion(add = .8)) +
  labs(y = "Proportion of captured individuals",
       fill = "Genus") +
  facet_grid(.~ landuse, switch = "x", space = "free_x", scales = "free_x",
             labeller = labeller(landuse = c("Food_forest" = "Food forest",
                                             "Forest" = "Forest",
                                             "Grassland" = "Grassland",
                                             "Arable_field" = "Cropland"))) +
  guides(fill = guide_legend(ncol = 2)) +
  set_theme + 
  theme(legend.position = "right")

# sum (should be almost same as mean)
tax_stack_sum <- dflist[["df_sitesum"]] %>%
  filter(count > 0) %>%
  left_join(codes_info) %>%
  ungroup() %>%
  mutate(landuse = factor(landuse, levels = c("Food_forest", "Forest",
                                              "Grassland", "Arable_field" ))) %>%
  # add taxonomic information
  left_join(taxonomy) %>%
  ggnested(aes(x = code, y = count, main_group = family, sub_group = genus), 
         main_palette = pal) +
  geom_col(position = position_fill(reverse = TRUE)) + 
  # 0 at level 0 of y axis
  scale_y_continuous(expand = expansion(mult = c(0, .05))) + #, labels = scales::percent
  # add padding around outer bars
  scale_x_discrete(expand = expansion(add = .8)) +
  labs(y = "Proportion of captured individuals",
       fill = "Genus") +
  facet_grid(.~ landuse, switch = "x", space = "free_x", scales = "free_x",
             labeller = labeller(landuse = c("Food_forest" = "Food forest",
                                             "Forest" = "Forest",
                                             "Grassland" = "Grassland",
                                             "Arable_field" = "Cropland"))) +
  guides(fill = guide_legend(ncol = 2)) +
  set_theme +
  theme(legend.position = "right")

tax_stack_mean
tax_stack_sum
```


## PERMDISP and PERMANOVA

### 3 betadispers

```{r PERMDISP}
# first for groups considered in main (pooled microarthropods and pooled macroarthropods)
# but also for separate groups

betadisper_output <- data.frame()
dmlist <- dmlist[grouporder_total]

for(m in names(dmlist)){
  set.seed(2983847)
  df <- as.matrix(dmlist[[m]])
  rn <- data.frame(code = rownames(df)) %>%
    left_join(codes_info %>% select(code, landuse, location))
  
  n_obs <- rn %>%
    group_by(landuse) %>%
    summarise(n = n()) %>%
    # now already remove land use type for which there are less than 6 observations
    filter(n >= 6)

  for(l in setdiff(n_obs$landuse, "Food_forest")){ # only for those reference land use types with >= 6 obs.
    dissim <- rn %>%
        filter(landuse %in% c("Food_forest",l)) %>%
        pull(code) %>%
      { df[.,.] } %>%
      as.dist()
    dissim_info <- rn %>% filter(landuse=="Food_forest"|landuse==l)
    disper_lu <- betadisper(dissim, dissim_info$landuse, type="centroid")
    FF <- round(disper_lu$group.distances[["Food_forest"]], digits=2)
    CON <- round(disper_lu$group.distances[[l]], digits=2)
    F_lu <- format(round(permutest(disper_lu)$tab$`F`[1], 
                         digits = 1))
    p_lu <- p.adjust(permutest(disper_lu)$tab$`Pr(>F)`[1],
                     method="holm", n = nrow(n_obs) - 1) # the number of comparisons made (FF vs ...)
    disper_loc <- betadisper(dissim, dissim_info$location, type="centroid")
    F_loc <- round(permutest(disper_loc)$tab$`F`[1], digits = 1)
    p_loc <- p.adjust(permutest(disper_loc)$tab$`Pr(>F)`[1], 
                      method="holm", n = nrow(n_obs) - 1) # idem above
    
    output <- cbind(group = m, 
                    contrast = l,
                    n_FF = n_obs %>% filter(landuse == "Food_forest") %>% pull(),
                    n_CON = n_obs %>% filter(landuse == l) %>% pull(),
                    F_loc, p_loc, FF, CON, F_lu, p_lu
                    )
    betadisper_output <- rbind(betadisper_output, output)
      }
}


flextable(betadisper_output)
write.csv(betadisper_output, "PERMDISP_output.csv")
```

Based on the results of the PERMDISP, we can perform all PERMANOVAs for the main groups. In a few cases there is the combination of a difference in the number of observations and dispersion between the two land use types, but then the dispersion is always greater in the group with the highest number of observations (FF), so there is a conservative bias, and yet we still detect differences, so OK; see supplement.

For the separate arthropod groups, there was never a statistically clear difference in dispersion between FF and the reference land use type, so all according separate PERMANOVAs can be conducted.

### 3 permanovas

```{r PERMANOVA}
permanova_output <- data.frame()

for(m in names(dmlist)){
  set.seed(2983847)
  df <- as.matrix(dmlist[[m]])
  rn <- data.frame(code = rownames(as.matrix(df))) %>%
    left_join(codes_info %>% select(code, landuse, location))
  
  n_obs <- rn %>%
    group_by(landuse) %>%
    summarise(n = n()) %>%
    # now already remove land use type for which there are less than 6 observations
    filter(n >= 6)

  for(l in setdiff(n_obs$landuse, "Food_forest")){ # only for those reference land use types with >= 6 obs.
      dissim <- rn %>%
          filter(landuse %in% c("Food_forest",l)) %>%
          pull(code) %>%
            { df[.,.] }
      dissim_info <- rn %>% filter(landuse=="Food_forest"|landuse==l)
      
      results <- adonis2(dissim ~ location + landuse, 
                          data = dissim_info, by="terms")
      
      results_loc <- results[1,] %>%
        rename("p" = "Pr(>F)",
               "pseudo_F" = "F") %>%
        mutate(df = paste0(Df, ", ", results[4,1]),
               p_adj = p.adjust(p, method = "holm", n = nrow(n_obs) - 1)) %>%
        select(df, pseudo_F, p_adj, R2)
      
      results_lu <- results[2,] %>%
        rename("p" = "Pr(>F)",
               "pseudo_F" = "F") %>%
        mutate(df = paste0(Df, ", ", results[4,1]),
               p_adj = p.adjust(p, method = "holm", n = nrow(n_obs) - 1)) %>%
        select(df, pseudo_F, p_adj, R2)
      
      output <- rbind(results_loc, results_lu) %>%
        rownames_to_column("term") %>%
        mutate(group = m,
               contrast = l,
               .before = "term")
      
      permanova_output <- rbind(permanova_output, round_df(output,digits=3))

    }
  
}
 
flextable(permanova_output)
write.csv(permanova_output, "PERMANOVA_output.csv")
```

