---
title: "DA_communitycomposition"
author: "Lieke"
date: "2024-06-20"
output: html_document
---

```{r packages}
# general
library(tidyverse)
# dissimilarity matrix, NMDS, PERMANOVA, PERMDISP
library(vegan)
#library(MicEco) #to calculate omega2
# models
library(glmmTMB)
library(mgcv) 
# visualisation
library(patchwork)
library(gratia) # for drawing contour plots from GAM
library(ggforce) # for NMDS amoebae
library(flextable)
library(gridExtra) # for grid.arrange
library(report)
library(phyloseq)
library(fantaxtic)
library(cowplot)
library(ggnested)
library(metR) # for geom_text_contour
landusetypes <- c("Food_forest", "Forest", "Grassland", "Arable_field")
```

```{r functions}
source("functions.R")

# for making the plot of ordisurf prettier in ggplot
extract.xyz <- function(obj) {
    xy <- expand.grid(x = obj$grid$x, y = obj$grid$y)
    xyz <- cbind(xy, c(obj$grid$z))
    names(xyz) <- c("x", "y", "z")
    return(xyz)
}
```

```{r data}
dmlist <- readRDS("dmlist.rds")
avgdissimlist <- readRDS("avgdissimlist.rds")

codes_info <- read.csv("Data/codes_info.csv")
sites_info <- codes_info %>%
  select(site, location, landuse) %>%
  distinct()

bblist_avgdissim <- readRDS("bblist_avgdissim.rds")

# read in data used for stacked bars taxonomy
taxon_macroarthropods <- readRDS("taxon_macroarthropods.rds")
taxon_earthworms <- readRDS("taxon_earthworms.rds")

# read in environmental (soil property) data
chemphys <- read.csv("Data/chemphys_P2.csv")

# read in NMDS results
NMDSreslist <- readRDS("NMDSreslist.rds")
```

```{r grouporder}
grouporder <- c("bacteria", "fungi", "protists", 
                "nematodes","microarthropods",  
                "macroarthropods", "earthworms")
                
grouporder_total <- c(grouporder,
                      "springtails", "mites",
                      "isopods", "millipedes","centipedes",
                      "harvestmen", "carabids")
```

# Hypothesis 3

## PERMDISP and PERMANOVA

### 3 betadispers

```{r PERMDISP}
# first for groups considered in main (pooled microarthropods and pooled macroarthropods)
# but also for separate groups

betadisper_output <- data.frame()
dmlist <- dmlist[grouporder_total]

for(m in names(dmlist)){
  set.seed(2983847)
  df <- as.matrix(dmlist[[m]])
  rn <- data.frame(code = rownames(df)) %>%
    left_join(codes_info %>% select(code, landuse, location))
  
  n_obs <- rn %>%
    group_by(landuse) %>%
    summarise(n = n()) %>%
    # now already remove land use type for which there are less than 6 observations
    filter(n >= 6)

  for(l in setdiff(n_obs$landuse, "Food_forest")){ # only for those reference land use types with >= 6 obs.
    dissim <- rn %>%
        filter(landuse %in% c("Food_forest",l)) %>%
        pull(code) %>%
      { df[.,.] } %>%
      as.dist()
    dissim_info <- rn %>% filter(landuse=="Food_forest"|landuse==l)
    disper_lu <- betadisper(dissim, dissim_info$landuse, type="centroid")
    FF <- round(disper_lu$group.distances[["Food_forest"]], digits=2)
    CON <- round(disper_lu$group.distances[[l]], digits=2)
    F_lu <- format(round(permutest(disper_lu)$tab$`F`[1], 
                         digits = 1))
    p_lu <- p.adjust(permutest(disper_lu)$tab$`Pr(>F)`[1],
                     method="holm", n = nrow(n_obs) - 1) # the number of comparisons made (FF vs ...)
    disper_loc <- betadisper(dissim, dissim_info$location, type="centroid")
    F_loc <- round(permutest(disper_loc)$tab$`F`[1], digits = 1)
    p_loc <- p.adjust(permutest(disper_loc)$tab$`Pr(>F)`[1], 
                      method="holm", n = nrow(n_obs) - 1) # idem above
    
    output <- cbind(group = m, 
                    contrast = l,
                    n_FF = n_obs %>% filter(landuse == "Food_forest") %>% pull(),
                    n_CON = n_obs %>% filter(landuse == l) %>% pull(),
                    F_loc, p_loc, FF, CON, F_lu, p_lu
                    )
    betadisper_output <- rbind(betadisper_output, output)
      }
}


flextable(betadisper_output)
write.csv(betadisper_output, "PERMDISP_output.csv")
```

Based on the results of the PERMDISP, we can perform all PERMANOVAs for the main groups. In a few cases there is the combination of a difference in the number of observations and dispersion between the two land use types, but then the dispersion is always greater in the group with the highest number of observations (FF), so there is a conservative bias, and yet we still detect differences, so OK; see supplement.

For the separate arthropod groups, there was never a statistically clear difference in dispersion between FF and the reference land use type, so all according separate PERMANOVAs can be conducted.

### 3 permanovas

```{r PERMANOVA}
permanova_output <- data.frame()

for(m in names(dmlist)){
  set.seed(2983847)
  df <- as.matrix(dmlist[[m]])
  rn <- data.frame(code = rownames(as.matrix(df))) %>%
    left_join(codes_info %>% select(code, landuse, location))
  
  n_obs <- rn %>%
    group_by(landuse) %>%
    summarise(n = n()) %>%
    # now already remove land use type for which there are less than 6 observations
    filter(n >= 6)

  for(l in setdiff(n_obs$landuse, "Food_forest")){ # only for those reference land use types with >= 6 obs.
      dissim <- rn %>%
          filter(landuse %in% c("Food_forest",l)) %>%
          pull(code) %>%
            { df[.,.] }
      dissim_info <- rn %>% filter(landuse=="Food_forest"|landuse==l)
      
      results <- adonis2(dissim ~ location + landuse, 
                          data = dissim_info, by="terms")
      
      results_loc <- results[1,] %>%
        rename("p" = "Pr(>F)",
               "pseudo_F" = "F") %>%
        mutate(df = paste0(Df, ", ", results[4,1]),
               p_adj = p.adjust(p, method = "holm", n = nrow(n_obs) - 1)) %>%
        select(df, pseudo_F, p_adj, R2)
      
      results_lu <- results[2,] %>%
        rename("p" = "Pr(>F)",
               "pseudo_F" = "F") %>%
        mutate(df = paste0(Df, ", ", results[4,1]),
               p_adj = p.adjust(p, method = "holm", n = nrow(n_obs) - 1)) %>%
        select(df, pseudo_F, p_adj, R2)
      
      output <- rbind(results_loc, results_lu) %>%
        rownames_to_column("term") %>%
        mutate(group = m,
               contrast = l,
               .before = "term")
      
      permanova_output <- rbind(permanova_output, round_df(output,digits=3))

    }
  
}
 
flextable(permanova_output)
write.csv(permanova_output, "PERMANOVA_output.csv")
```

## NMDS

```{r NMDS}
# list to save plots in
NMDS_plots <- list()

# list to save results in
NMDSreslist <- list() # will be chill to have this saved instead of having to rerun nmds all the time

for(m in names(dmlist[grouporder])){
  df <- dmlist[[m]]
  set.seed(2983847)
  NMDS_res <- metaMDS(df, k=2, trymax = 999, trace = FALSE) 
  # autotransform by default TRUE
  
  p <- myNMDS(df, NMDS_res, codes_info) + 
    # temporary for figure
    labs(x = "NMDS1", y = "NMDS2")
  
  NMDS_plots[[m]] <- p
  
  NMDSreslist[[m]] <- NMDS_res

}

saveRDS(NMDSreslist, "NMDSreslist.rds")
# (NMDSplot <- do.call("grid.arrange", c(NMDS_plots, ncol = 3)))
(NMDSplot <- do.call("plot_grid", c(NMDS_plots, ncol=3, align="hv")))
ggsave("NMDSplot_height17.png", NMDSplot, width = 16, height = 16.5)
ggsave("NMDSplot_height17.svg", NMDSplot, width = 16, height = 17)

## separately for separate arthropod groups
## PERMANOVA only if at least 6 observations per land use type > for now also only these on NMDS
## comparison with forests possible for all groups
## comparison with grassland only possible for springtails, mites, isopods and carabids
## comparison with arable only possible for springtails and carabids

NMDS_plots_sep <- list()

for(m in c("springtails", "mites",
           "isopods", "centipedes", "millipedes",
           "harvestmen", "carabids")){
  set.seed(2983847)
  df <- as.matrix(dmlist[[m]])

  rn <- data.frame(code = rownames(df)) %>%
    left_join(codes_info %>% select(code, landuse, location))
  
  # count number of observations per land use type
  n_obs <- rn %>%
    group_by(landuse) %>%
    summarise(n = n()) %>%
    # now already remove land use type for which there are less than 6 observations
    filter(n >= 6)
  
  # filter out rownames (and column names) of land use type with too few observations
  rn <- rn %>%
    filter(landuse %in% n_obs$landuse)
  
  df <- df[rn$code, rn$code]

    
  p <- myNMDS(df, codes_info) + 
    # temporary for figure
    labs(x = "NMDS1", y = "NMDS2")
  
  NMDS_plots_sep[[m]] <- p

}



(NMDSplot_sep <- do.call("grid.arrange", c(NMDS_plots_sep, ncol = 3)))
# ggsave("NMDSplot_sep.svg", NMDSplot_sep, width = 16, height = 16)
```

```{r legend}
legend <- ggpubr::get_legend(p)
as_ggplot(legend)
```

```{r NMDS fit}
#Check the stress and fit of the nematodes NMDS following https://uw.pressbooks.pub/appliedmultivariatestatistics/chapter/nmds/
df <- dmlist[["nematodes"]]
NMDS_res <- metaMDS(df, k=2, trymax = 999, trace = FALSE) 
fit <- goodness(NMDS_res)
NMDS_res$stress
plot(NMDS_res, display = "sites", type = "none")
points(NMDS_res, display = "sites", cex = 2*fit/mean(fit)) #for nematodes mostly driven by few samples?
stressplot(NMDS_res) #Shephard plot..large scatter indicates issues
```

## ordisurf & GAM

Here, we use ordisurf to explore potential relationships between key soil properties and variation in soil community composition as represented in the ordination.
We evaluate a) if the soil property under consideration displays a statistically clear (p < 0.05) relationship with the ordination (the data scores of the plots); b) if this relationship is sufficiently close to a linear relationship so that it would be acceptable to plot it as a vector on the NMDS (which can be done using the envfit function), or if the relationship is not linear and we should rather show the plot generated by ordisurf (a contour plot).

Note that: "The fitted GAM is a regression model and has the usual assumptions of such models. Of particular note is the assumption of independence of residuals. If the observations are not independent (e.g. they are repeat measures on a set of objects, or from an experimental design, inter alia) do not trust the p-values from the GAM output."
In our case, this means that it is suboptimal to not include the location random effect here (spatial correlation).
Therefore, we also make our own GAM, in which we include the location random effect

also see https://fromthebottomoftheheap.net/2011/06/10/what-is-ordisurf-doing/
and https://fromthebottomoftheheap.net/2021/02/02/random-effects-in-gams/

```{r NMDS ordisurf_GAM}
# df for model summaries
modeloutput_ordisurf_GAM <- data.frame()

# plotlist for contour plots 
ordisurfplots <- list()
gamplots <- list()

# for plotting, values for each of five soil properties to use as binwidth
binwidths <- data.frame(group = rep(c("bacteria", "fungi", "protists",
                                  "nematodes","microarthropods", "earthworms"), each = 5),
                        property = rep(c("bulkdensity", "pH_KCl", "C_total", "N_total", "P_total"), times = 6),
                        bin_g = c(rep(c(0.05, 0.2, 0.1, 0.01, 100), times = 5), # can play around with this a bit still
                                  c(0.02,0.01,0.05,0.005,10))) # different for earthworms at least

# cleanly collect output from ordisurf and GAM runs - not for macroarthropods
for(m in c("bacteria", "fungi", "protists", 
            "nematodes","microarthropods",  
            "earthworms")){
  df <- dmlist[[m]]
  
  NMDS_res <- NMDSreslist[[m]]
  
  scores <- data.frame(scores(NMDS_res, display = "sites", scaling = 3)) 
    
  env <- data.frame(code = rownames(scores)) %>%
    left_join(chemphys %>% select(code, location, landuse,
                                  bulkdensity, pH_KCl, C_total, N_total, P_total), by = "code") %>%
    mutate(location = factor(location),
           landuse = factor(landuse, levels = c("Food_forest", "Forest",
                                                "Grassland", "Arable_field"))) %>%
    cbind(scores) %>% # this is OK, correct, codes match rownames, so correct environmental data bound with scores
    # make longer to enable loop below - but then rownames disappear, so keep column code and move to rownames in loop
    pivot_longer(cols = c(bulkdensity, pH_KCl, C_total, N_total, P_total),
                 names_to = "property")

  # fit and add 2d surface using ordisurf (simple GAM) and fit GAM with location RE for each of the separate soil properties
  for(g in c("bulkdensity", "pH_KCl", "C_total", "N_total", "P_total")){
    
    # pick soil property
    env_g <- env %>%
      filter(property == g) %>%
      #column_to_rownames(var = "code") %>%
      mutate(z = NA) # necessary later on for plotting
    
    # pick conditional binwidth
    bin_g <- binwidths %>% filter(group == m & property == g) %>% pull(bin_g)
    
    # ordisurf
    surf <- ordisurf(NMDS_res ~ value, data = env_g, 
                     method = "REML", 
                     select = TRUE) 
    
    # GAM with location RE
    gam_RE <- gam(value ~ s(NMDS1, NMDS2, k=10, bs="tp", fx = FALSE) + s(location, bs = "re"), 
                  data = env_g, 
                  family = Gamma(link = "log"),
                  method = "REML",
                  select = TRUE)
    
    # save output we just select the p-values from the output of the gam here
    output <- data.frame(group = m,
                         property = g,
                         p_ordisurf = round(summary(surf)$s.table[,4], 3), # this extracts the p-value (fourth column)
                         p_GAM_RE = round(summary(gam_RE)$s.table[1,4], 3), # this extracts p-value for s(NMDS1, NMDS2)
                         p_GAM_RE_location = round(summary(gam_RE)$s.table[2,4], 3),
                         R2_adj_GAM_RE = round(summary(gam_RE)$r.sq,3)) 
    
    modeloutput_ordisurf_GAM <- rbind(modeloutput_ordisurf_GAM, output)
    
    # make pretty plot for GAM with RE
    
    ## predict with tidygam
    x2 <- as.data.frame(predict_gam(gam_RE, length_out = 50, tran_fun = exp)) %>% 
      # higher length_out is smoother contours
      rename(z = names(.)[4]) %>%
      mutate(z = as.vector(z)) # for some reason this was a matrix instead of a vector, hence the [ ,1]
    
    ## plot
    p <- ggplot() + 
      geom_point(data = env_g, aes(x = NMDS1, y = NMDS2, colour = landuse)) + 
      scale_colour_manual(values = c("#56B4E9", "#D55E00","#009E73", "#F0E442")) +
      stat_contour(data = x2, aes(x=NMDS1, y=NMDS2, z=z), binwidth = bin_g) + # maybe make lines a bit thicker still 
      geom_text_contour(data = x2, aes(x=NMDS1, y=NMDS2, z=z), vjust=+2, 
                        check_overlap = TRUE, size = 3,
                        skip = 0, binwidth = bin_g) + # may want to skip = 1 if we have many lines
      coord_equal() +
        labs(#title = g,
             x = "NMDS1", y = "NMDS2") + 
        theme_minimal(base_size = 16) +
        theme(legend.position = "none",
              panel.grid.major = element_line(size = 0.5), 
              panel.grid.minor = element_line(size = 0.5),
              panel.border = element_rect(colour = "black", 
                                          fill = NA, size = 0.5))
     
    gamplots[[m]][[g]] <- p

  }

} 

# write.csv(modeloutput_ordisurf_GAM, "modeloutput_ordisurf_GAM.csv")
# saveRDS(ordisurfplots, "ordisurfplots.rds")
# saveRDS(gamplots, "gamplots.rds")
p1 <- wrap_plots(gamplots[["bacteria"]], ncol = 5)
p2 <- wrap_plots(gamplots[["fungi"]], ncol = 5)
p3 <- wrap_plots(gamplots[["protists"]], ncol = 5)
p4 <- wrap_plots(gamplots[["nematodes"]], ncol = 5)
p5 <- wrap_plots(gamplots[["microarthropods"]], ncol = 5)
p6 <- wrap_plots(gamplots[["earthworms"]], ncol = 5) # pH for now doesn't want to show

pup <- wrap_plots(p1, p2, p3, nrow = 3)
plow <- wrap_plots(p4, p5, p6, nrow = 3)
ggsave("gamplots_up.png", pup, width = 500, height = 350, unit ="mm", dpi = 100) # should later be 600
ggsave("gamplots_low.png", plow, width = 500, height = 350, unit ="mm", dpi = 100)

# test to look into cutoff for visualisation
test <- modeloutput_ordisurf_GAM %>%
  filter(p_GAM_RE < 0.05 & R2_adj_GAM_RE > 0.6)

# OK, so now almost all are significant
# saveRDS(ordisurfplots, "ordisurfplots.rds")
# ordisurfplots <- readRDS("ordisurfplots.rds")
# compare ordisurf plots and gam with re plots

```

## Taxonomy

```{r taxonomy_bars}
ps_bact <- readRDS("Phyloseq_files/ps_bacteria.RDS")
ps_non_am_fun <- readRDS("Phyloseq_files/ps_nonamfungi.RDS")
ps_pro <- readRDS("Phyloseq_files/ps_protists.RDS")
ps_nema <- readRDS("Phyloseq_files/ps_nematodes.RDS")
ps_microarthr <- readRDS("Phyloseq_files/ps_microarthropods.RDS")
taxon_macroarthropods <- readRDS("taxon_macroarthropods.rds")
taxon_earthworms <- readRDS("taxon_earthworms.rds")
more_site_info <- read_csv2("Phyloseq_files/more_site_info.csv")

#colors
# tol10qualitative=c("#332288", "#88CCEE", "#44AA99", "#117733", "#999933", "#DDCC77", "#661100", "#CC6677", "#882255", "#AA4499")
test <- c("#0C7BDC", "#FFC20A","#40B0A6","#5D3A9B","#FEFE62","#DC3220","#1AFF1A","#D35FB7","#994F00","#88CCEE")

#theme
set_theme <- list(
  theme_bw(),
  theme(panel.grid = element_blank(),
        strip.background.x = element_blank(), 
        strip.placement = "outside",
        axis.line = element_line(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.length.x = unit(0, "pt"),
        panel.border = element_blank(),
        panel.spacing.x = unit(7, "pt"), # does what it implies, space the panels of the facet 
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        strip.text.x = element_text(size = 14))#,
        # legend.text = element_text(size = 13), 
        # legend.title = element_text(size = 14))
)
```

```{r taxonomy_bact}
#Bacteria
sdf <- data.frame(sample_data(ps_bact))
sdf <- sdf %>% left_join(select(more_site_info, code, site), by = "code")
sample_data(ps_bact)$unique_site <- sdf$site
sample_data(ps_bact)$unique_site <- as.factor(sample_data(ps_bact)$unique_site)

#site metadata
site_meta <- sample_data(ps_bact) %>%
  as.data.frame() %>%
  group_by(unique_site) %>%
  summarise_all(~ unique(.)[1]) %>%
  select(unique_site, landuse)
site_meta <- sample_data(site_meta)
rownames(site_meta) <- site_meta$unique_site

#merge per site
ps_site <- merge_samples(ps_bact, "unique_site", fun=mean)
sample_data(ps_site) <- sample_data(site_meta)
sample_data(ps_site)$landuse <- factor(sample_data(ps_site)$landuse,
                                       levels=c("Food_forest","Forest","Grassland","Arable_field"))
top_nested <- nested_top_taxa(ps_site,
                              top_tax_level = "Phylum",
                              nested_tax_level = "Class",
                              n_top_taxa = 10, 
                              n_nested_taxa = 3)
(nb <- plot_nested_bar(ps_obj = top_nested$ps_obj,
                top_level = "Phylum",
                nested_level = "Class",
                palette=test,
                legend_title = "Phylum and classes") +
  scale_x_discrete(expand = expansion(add = .8)) +
  facet_grid(.~ landuse, switch = "x", space = "free_x", scales = "free_x",
             labeller = labeller(landuse = c("Food_forest" = "Food forest",
                                             "Forest" = "Forest",
                                             "Grassland" = "Grassland",
                                             "Arable_field" = "Cropland"))) +
  labs(x = "Land use", y = "Mean relative abundance") +
  guides(fill = guide_legend(ncol = 2)) +
  set_theme + theme(legend.position = "right"))
ggsave("bact_tax.png",nb, width = 300, height = 200, unit="mm", dpi=600)
# leg <- get_legend(nb)
# (cow <- cowplot::plot_grid(nb_no_leg, leg, ncol=2, rel_widths = c(2, 1)))
# ggsave("bact_try.png", cow, width = 300, height = 150, unit="mm", dpi=600)
```

```{r taxonomy_fun}
#Non-AM fungi
sdf <- data.frame(sample_data(ps_non_am_fun))
sdf <- sdf %>% left_join(select(more_site_info, code, site), by = "code")
sample_data(ps_non_am_fun)$unique_site <- sdf$site
sample_data(ps_non_am_fun)$unique_site <- as.factor(sample_data(ps_non_am_fun)$unique_site)

#site metadata
site_meta <- sample_data(ps_non_am_fun) %>%
  as.data.frame() %>%
  group_by(unique_site) %>%
  summarise_all(~ unique(.)[1]) %>%
  select(unique_site, landuse)
site_meta <- sample_data(site_meta)
rownames(site_meta) <- site_meta$unique_site

#merge per site
ps_site <- merge_samples(ps_non_am_fun, "unique_site", fun=mean)
sample_data(ps_site) <- sample_data(site_meta)
sample_data(ps_site)$landuse <- factor(sample_data(ps_site)$landuse,
                                       levels=c("Food_forest","Forest","Grassland","Arable_field"))
top_nested <- nested_top_taxa(ps_site,
                              top_tax_level = "Phylum",
                              nested_tax_level = "Class",
                              n_top_taxa = 10, 
                              n_nested_taxa = 3)
(nb <- plot_nested_bar(ps_obj = top_nested$ps_obj,
                top_level = "Phylum",
                nested_level = "Class",
                palette=test,
                legend_title = "Phylum and classes") +
  scale_x_discrete(expand = expansion(add = .8)) +
  facet_grid(.~ landuse, switch = "x", space = "free_x", scales = "free_x",
             labeller = labeller(landuse = c("Food_forest" = "Food forest",
                                             "Forest" = "Forest",
                                             "Grassland" = "Grassland",
                                             "Arable_field" = "Cropland"))) +
  labs(y = "Mean relative abundance") +
  guides(fill = guide_legend(ncol = 2)) +
  set_theme + theme(legend.position = "right"))
ggsave("fun_tax.png",nb, width = 300, height = 200, unit="mm", dpi=600)
# leg <- get_legend(nb)
# (cow <- cowplot::plot_grid(nb_no_leg, leg, ncol=2, rel_widths = c(2, 1)))
# ggsave("fun_try.png", cow, width = 400, height = 150, unit="mm", dpi=600)
```

```{r taxonomy_pro}
#Protists
sdf <- data.frame(sample_data(ps_pro))
sdf <- sdf %>% left_join(select(more_site_info, code, site), by = "code")
sample_data(ps_pro)$unique_site <- sdf$site
sample_data(ps_pro)$unique_site <- as.factor(sample_data(ps_pro)$unique_site)

#site metadata
site_meta <- sample_data(ps_pro) %>%
  as.data.frame() %>%
  group_by(unique_site) %>%
  summarise_all(~ unique(.)[1]) %>%
  select(unique_site, landuse)
site_meta <- sample_data(site_meta)
rownames(site_meta) <- site_meta$unique_site

#merge per site
ps_site <- merge_samples(ps_pro, "unique_site", fun=mean)
sample_data(ps_site) <- sample_data(site_meta)
sample_data(ps_site)$landuse <- factor(sample_data(ps_site)$landuse,
                                       levels=c("Food_forest","Forest","Grassland","Arable_field"))
ps_site %>% rank_names()
top_nested <- nested_top_taxa(ps_site,
                              top_tax_level = "Subdivision",
                              nested_tax_level = "Class",
                              n_top_taxa = 10, 
                              n_nested_taxa = 3)
(nb <- plot_nested_bar(ps_obj = top_nested$ps_obj,
                top_level = "Subdivision",
                nested_level = "Class",
                palette=test,
                legend_title = "Subdivision and classes") +
  scale_x_discrete(expand = expansion(add = .8)) +
  facet_grid(.~ landuse, switch = "x", space = "free_x", scales = "free_x",
             labeller = labeller(landuse = c("Food_forest" = "Food forest",
                                             "Forest" = "Forest",
                                             "Grassland" = "Grassland",
                                             "Arable_field" = "Cropland"))) +
  labs(y = "Mean relative abundance") +
  guides(fill = guide_legend(ncol = 2)) +
  set_theme + theme(legend.position = "right"))
ggsave("protist_tax.png",nb, width = 300, height = 200, unit="mm", dpi=600)
```

```{r taxonomy_nema}
#nematodes
sdf <- data.frame(sample_data(ps_nema))
sdf <- sdf %>% left_join(select(more_site_info, code, site), by = "code")
sample_data(ps_nema)$unique_site <- sdf$site
sample_data(ps_nema)$unique_site <- as.factor(sample_data(ps_nema)$unique_site)

#site metadata
site_meta <- sample_data(ps_nema) %>%
  as.data.frame() %>%
  group_by(unique_site) %>%
  summarise_all(~ unique(.)[1]) %>%
  select(unique_site, landuse)
site_meta <- sample_data(site_meta)
rownames(site_meta) <- site_meta$unique_site

#merge per site
ps_site <- merge_samples(ps_nema, "unique_site", fun=mean)
sample_data(ps_site) <- sample_data(site_meta)
sample_data(ps_site)$landuse <- factor(sample_data(ps_site)$landuse,
                                       levels=c("Food_forest","Forest","Grassland","Arable_field"))
top_nested <- nested_top_taxa(ps_site,
                              top_tax_level = "Class",
                              nested_tax_level = "Family",
                              n_top_taxa = 10, 
                              n_nested_taxa = 3)
(nb <- plot_nested_bar(ps_obj = top_nested$ps_obj,
                top_level = "Class",
                nested_level = "Family",
                palette=test,
                legend_title = "Class and families") +
      scale_x_discrete(expand = expansion(add = .8)) +
      facet_grid(.~ landuse, switch = "x", space = "free_x", scales = "free_x",
             labeller = labeller(landuse = c("Food_forest" = "Food forest",
                                             "Forest" = "Forest",
                                             "Grassland" = "Grassland",
                                             "Arable_field" = "Cropland"))) +
  labs(y = "Mean relative abundance") +
  guides(fill = guide_legend(ncol = 2)) +
  set_theme + theme(legend.position = "right"))
ggsave("nema_tax.png",nb, width = 300, height = 200, unit="mm", dpi=600)
```

```{r taxonomy_microarthr}
#Microarthropods
sdf <- data.frame(sample_data(ps_microarthr))
sdf <- sdf %>% rename(code = vb_code)
sdf <- sdf %>% left_join(select(more_site_info, code, site), by = "code")
sample_data(ps_microarthr)$unique_site <- sdf$site
sample_data(ps_microarthr)$unique_site <- as.factor(sample_data(ps_microarthr)$unique_site)

#site metadata
site_meta <- sample_data(ps_microarthr) %>%
  as.data.frame() %>%
  group_by(unique_site) %>%
  summarise_all(~ unique(.)[1]) %>%
  select(unique_site, landuse)
site_meta <- sample_data(site_meta)
rownames(site_meta) <- site_meta$unique_site

#merge per site
ps_site <- merge_samples(ps_microarthr, "unique_site", fun=mean)
sample_data(ps_site) <- sample_data(site_meta)
sample_data(ps_site)$landuse <- factor(sample_data(ps_site)$landuse,
                                       levels=c("Food_forest","Forest","Grassland","Arable_field"))
top_nested <- nested_top_taxa(ps_site,
                              top_tax_level = "Class",
                              nested_tax_level = "Family",
                              n_top_taxa = 10, 
                              n_nested_taxa = 3)
# top_nested$ps_obj@tax_table[,"Order"] <- factor(top_nested$ps_obj@tax_table[,"Order"], 
#                                                    levels=c("Other", "Mesostigmata","Sarcoptiformes",
#                                                             "Trombidiformes", "Entomobryomorpha", "Neelipleona",
#                                                             "Poduromorpha", "Symphypleona"))
(nb <- plot_nested_bar(ps_obj = top_nested$ps_obj,
                top_level = "Class",
                nested_level = "Family",
                palette=test,
                legend_title="Class and families"
                ) +
   scale_x_discrete(expand = expansion(add = .8)) +
   facet_grid(.~ landuse, switch = "x", space = "free_x", scales = "free_x",
             labeller = labeller(landuse = c("Food_forest" = "Food forest",
                                             "Forest" = "Forest",
                                             "Grassland" = "Grassland",
                                             "Arable_field" = "Cropland"))) +
  labs(y = "Mean relative abundance") +
  guides(fill = guide_legend(ncol = 2)) +
  set_theme + theme(legend.position = "right"))
ggsave("microarthr_tax.png",nb, width = 300, height = 200, unit="mm", dpi=600)
```

```{r taxonomy_macroarthr}
#macroarthropods
(nb <- taxon_macroarthropods %>% 
  left_join(sites_info) %>%
  mutate(landuse = factor(landuse, levels = c("Food_forest", "Forest",
                                              "Grassland", "Arable_field"))) %>%
 ggnested(aes(x = site, y = mean_quant, main_group = class, sub_group = family),
         main_palette=test) +
  geom_col(position = position_fill(reverse = TRUE)) + # colour = "black"
  # 0 at level 0 of y axis
  scale_y_continuous(expand = expansion(mult = c(0, .05))) + #, labels = scales::percent
  # add padding around outer bars
  scale_x_discrete(expand = expansion(add = .8)) +
  labs(y = "Proportion of captured individuals",
       fill = "Genus") +
  facet_grid(.~ landuse, switch = "x", space = "free_x", scales = "free_x",
             labeller = labeller(landuse = c("Food_forest" = "Food forest",
                                             "Forest" = "Forest",
                                             "Grassland" = "Grassland",
                                             "Arable_field" = "Cropland"))) +
  guides(fill = guide_legend(ncol = 2)) +
  set_theme + theme(legend.position = "right"))
ggsave("macroarthr_tax.png",nb, width = 300, height = 200, unit="mm", dpi=600)
```

```{r taxonomy_ew}
#earthworms
(nb <- taxon_earthworms %>% 
  left_join(sites_info) %>%
  mutate(landuse = factor(landuse, levels = c("Food_forest", "Forest",
                                              "Grassland", "Arable_field"))) %>%
  # filter out empty samples, otherwise get empty bars
  filter(mean_quant > 0) %>%
ggnested(aes(x = site, y = mean_quant, main_group = genus, sub_group = species),
         main_palette=test) +
  geom_col(position = position_fill(reverse = TRUE)) + # colour = "black"
  # 0 at level 0 of y axis
  scale_y_continuous(expand = expansion(mult = c(0, .05))) + #, labels = scales::percent
  # add padding around outer bars
  scale_x_discrete(expand = expansion(add = .8)) +
  labs(y = "Proportion of captured individuals")+
  facet_grid(.~ landuse, switch = "x", space = "free_x", scales = "free_x",
             labeller = labeller(landuse = c("Food_forest" = "Food forest",
                                             "Forest" = "Forest",
                                             "Grassland" = "Grassland",
                                             "Arable_field" = "Cropland"))) +
  guides(fill = guide_legend(ncol = 2)) +
  set_theme + theme(legend.position = "right"))
ggsave("ew_tax.png",nb, width = 300, height = 200, unit="mm", dpi=600)
```


## Indicator species analysis? 
For now we're not including this for the paper, but it is easy and fun, gives interesting results that make the findings a bit more tangible and ecologically interpretable and hint at the importance of looking at the identity and ecology of the species.

# Hypothesis 4

So, in this section of the results, we only look at within-site dissimilarity.
We look at average pairwise dissimilarity rather than multisite dissimilarity. This is due to methodological constraints, because we wanted to use "rarefied dissimilarities" for the sequencing data. Thus, we here use the dissimilarity matrices that have been produced (or read) upstream and have been used for NDMS, PERMANOVA and PERMDISP.

See "Model_diagnostics_community_composition.Rmd" for the checks of the model assumptions.
The bootstrapping is done as for other models (see "bootstrap_approach.R").

As the Bray-Curtis dissimilarity is bound between 0 and 1, we use a beta regression. However, in some cases this dissimilarity is equal to 0 or 1. We chose to respond to this by using a transformation of the dissimilarity following Smithson and ... (2006). We did consider using an approach for a mixture distribute, with three separate models, but ultimately deemed this approach of transformation more useful and perhaps not that problematic (because perhaps no qualitative difference between dissimilarity of 1 and 0.99 in this case where we look at within-site dissimilarity and could in fact expect it to be possible that there would be at least some species (e.g. of these mobile species groups) occurring in both traps/more than 1 plot)

## Plotting

```{r avgdissim_plot}
# list for storing plots and dataframe for storing total model summary
avgdissim_plots <- list()
avgdissim_modelsumm <- data.frame()

# dataframe for storing estimates of dispersion coefficients
dispcoeff <- data.frame()

for (m in c("bacteria", "fungi", "protists",
            "nematodes",
            "microarthropods",
            "macroarthropods",
            "earthworms")){ # not doing it for separate microarthropod and macroarthropod groups
  
  # Prepare dataframe
  ## Transform to fit (0,1) (barely changes the values) and make landuse factor
  df <- avgdissimlist[[m]] %>%
      mutate(avgdissim_transf = (avg_pairwise_dissim * (length(avg_pairwise_dissim) - 1)
                                    + 0.5)/length(avg_pairwise_dissim),
             landuse = factor(landuse, levels = c("Food_forest", "Forest", 
                                                "Grassland", "Arable_field"))) %>%
  ## Removing land use type for which there are less than 6 observations 
      na.omit() %>%
      group_by(landuse) %>%
      filter(n() >= 6) %>%
      ungroup()

  # model
  fitted_model <- glmmTMB(avgdissim_transf ~ landuse,
                     family = beta_family(),
                     data = df)
  
  
  # put dispersion coefficient in dataframe
  dispcoeff <- rbind(dispcoeff, data.frame(Group = m,
                                           Dispersion = round(sigma(fitted_model), 
                                                              digits = 2)))
  # process bootstrap values 
  # to obtain 95%CIs of means response scale and p-values
  bb <- bblist_avgdissim[[m]]
  
  ## make dataframe from bootstrap estimates, 
  ## with landuse means and differences in logit scale
  bb_df_logit <- as.data.frame(bb[["t"]]) %>%
    # rename columns
    rename("Food_forest_mean" = "(Intercept)") %>%
    rename_with(.fn = ~ gsub("landuse", "diff_",.x), starts_with("landuse")) %>% 
    # add column with means of reference landuses, by adding their estimated difference to the estimated mean of the food forest
    # some land use types were excluded, so takes this into account
    mutate(Forest_mean = if("diff_Forest" %in% names(.)) {
      Food_forest_mean + diff_Forest} else NULL,
           Grassland_mean = if("diff_Grassland" %in% names(.)) {
             Food_forest_mean + diff_Grassland} else NULL,
           Arable_field_mean = if("diff_Arable_field" %in% names(.)) {
             Food_forest_mean + diff_Arable_field} else NULL,
           .after = "Food_forest_mean")

  ## make dataframe with landuse means in response scale
  bb_df_response <- bb_df_logit %>%
    select(ends_with("mean")) %>%
    # fist reverse the Smithson and Verkuilen transformation
    mutate(across(everything(), ~ ((.) * length(df$avg_pairwise_dissim) - 0.5)/(length(df$avg_pairwise_dissim) - 1))) %>%
    # back-transform from logit
    mutate(across(everything(), ~ exp(.) / (1 + exp(.))))

  
  ## obtain CIs for model estimated means for plotting
  ## we plot the model estimated mean and make the CI around this by using the SD obtained from the parametric bootstrapping 
  
  mean_boot_sd <- bb_df_response %>%
    pivot_longer(names_to = "landuse", names_pattern = "(.*)_mean",
                 cols = ends_with("mean")) %>%
    group_by(landuse) %>%
    reframe(boot_sd = sd(value))
  
  estimates_response <- data.frame(model_estimate = fixef(fitted_model)[["cond"]]) %>% 
    rownames_to_column(var = "landuse") %>%
    # change landuse names
    mutate(landuse = case_when(landuse == "(Intercept)" ~ "Food_forest",
                             .default = str_remove(landuse, "^landuse"))) %>%
    # compute actual means for reference land uses (i.e. subtract from food forest mean)
    mutate(model_mean = ifelse(landuse != "Food_forest", 
                             model_estimate[landuse == "Food_forest"] + model_estimate,
                             model_estimate)) %>%
    # reverse Smithson and Verkuilen transformation
    mutate(across(where(is.numeric), ~ ((.) * length(df$avg_pairwise_dissim) - 0.5)/(length(df$avg_pairwise_dissim) - 1))) %>%
    # back-transform to response scale
    mutate(across(where(is.numeric), ~ exp(.) / (1 + exp(.)))) %>%
    # join with sd values
    left_join(mean_boot_sd) %>%
    # add CI limits
    mutate(lower.CL = model_mean - 1.96*boot_sd,
           upper.CL = model_mean + 1.96*boot_sd) %>%
    # format for ggplot_the_response
    select(landuse, model_mean, lower.CL, upper.CL) %>%
    rename(response = model_mean)
    
  ## obtain two-tailed p-values, 
  boot_pval <- function(x) { 2*min(mean(x<0), mean(x>0)) } 
  boot_contrasts <- data.frame(p_raw = apply(bb$t[, colnames(bb$t) != "(Intercept)"], 
                                             2, boot_pval)) %>%
    # adjust p-values for multiple comparisons using Holm's method (FWER) 
    mutate(p_adj_Holm = p.adjust(p_raw, method = "holm"))  %>% 
    # already format to supply to ggplot_the_response
    # adding column contrast
    rownames_to_column(var = "contrast") %>%
    mutate(contrast = paste0(str_remove(contrast, "^landuse"), " - Food_forest")) %>%
    rename(p.value = p_adj_Holm)

  
  ## plotting (the difference between original values and those after Smithson & Verkuilen transformation is minute)
  p <- ggplot_the_response(
    beta = TRUE,
    fit = fitted_model,
    fit_emm = estimates_response, 
    fit_pairs = boot_contrasts, 
        dots = "sina1",
      contrast_rows = "all",
      palette = pal_okabe_ito_FF,
      y_label = "Average pairwise Bray-Curtis dissimilarity",
      g_label = "none"
    ) + 
      theme(axis.text.x = element_blank(),
            axis.title.y = element_text(size = 18),
            axis.text.y = element_text(size = 15),
            axis.ticks.x = element_blank()) +     
    # make sure y-axis labels are always round numbers, except in case of springtails 1 digit
      scale_y_continuous(labels = if(m == "springtails"){function(x) sprintf("%.1f", x)} 
                         else function(x) sprintf("%.0f", x)) +
      # code to add values for land use type that was not included in model due to too few values
      {if(!(all(landusetypes %in% df$landuse)))
      geom_point(data = avgdissimlist[[m]] %>% 
                   # add the values for the landuse type that is missing in the df
                   filter(landuse == landusetypes[!(landusetypes %in% df$landuse)]), 
                 aes(x = landuse, y = avg_pairwise_dissim), size = 3.5, alpha = 0.6)} +
      # need to overwrite some plot scaling
      coord_cartesian(ylim = c(0, 1.35)) + # otherwise padding below 0.0
      scale_y_continuous(expand = c(0,0), limits = c(0,1.35), breaks = seq(0, 1, by = 0.2))
  
  
  avgdissim_plots[[m]] <- p
  
}

# (avgdissimplot <- do.call("grid.arrange", c(avgdissim_plots, ncol = 3)))
avgdissimplot <- wrap_plots(avgdissim_plots, ncol = 3) # to get x- and y-axes aligned
ggsave("avgdissimplot.png", avgdissimplot, width = 360, height = 320, unit="mm", dpi=600)
ggsave("avgdissimplot.svg", avgdissimplot, width = 360, height = 320, unit="mm", dpi=600)
ggsave("avgdissimplot_legend.svg", avgdissimplot$bacteria + theme(legend.position = "bottom"), width = 100, height = 100, unit="mm", dpi=600)

write.csv(dispcoeff, "beta_dispersioncoeff.csv")
```

## model summary 
Not reporting R² maybe?
https://stats.stackexchange.com/questions/552281/calculating-different-pseudo-r2-for-a-betareg-model
Achim Zeileis: "I generally find the pseudo-R-squared to be of rather limited use in beta regression. Whether the model fit is useful (for you) depends on many aspects, e.g., whether you mostly want a model for the mean or you are really interested in a probabilistic fit. A comprehensive answer for this is beyond the scope of this question, though. Alternative means of model comparisons could include information criteria (AIC or BIC), scores for the mean predicitons like (root) mean-squared error (MSE, RMSE), or probabilistic scoring rules (CRPS, log-score), or graphical model assessments (quantile-quantile plots, PIT histograms, etc.)."

For now not showing Fit-measures, because unsure how useful they would be and did not sufficiently read up on different options
not showing CIs or p-values because they are already shown in figure from bootstrapping approach and these from model summary wouldn't correspond
But then what is left to show? Model estimated means are also plotted already.
That leaves the dispersion coefficient, which we could perhaps just add in the subscript of the figure

