---
title: "DA_communitycomposition"
author: "Lieke"
date: "2024-06-20"
output: html_document
---

```{r packages}
# general
library(tidyverse)
# dissimilarity matrix, NMDS, PERMANOVA, PERMDISP
library(vegan)
# models
library(glmmTMB)
library(mgcv) 
# indicator species
library(indicspecies)
# visualisation
library(patchwork)
library(ggforce) # for NMDS amoebae
library(flextable)
library(gridExtra) # for grid.arrange
library(report)
library(fantaxtic)
library(cowplot)
library(ggnested)
landusetypes <- c("Food_forest", "Forest", "Grassland", "Arable_field")
```

```{r functions}
source("functions.R")
```

```{r data}

# original data
harvestmen_raw <- read.csv("Data/Harvestmen.csv") %>%
  select(!c(Stage, Sex)) %>%
  as.data.frame() %>%
  # change from all caps to normal
  mutate(species = str_to_title(species))

dflist <- readRDS("dflist.rds")

# species abundance matrices (community data table)
specablist <- readRDS("specablist.rds")

# Bray-Curtis dissimilarity
dissimlist <- readRDS("dissimlist.rds")

# code info
codes_info <- read.csv("Data/codes_info.csv")

# taxonomic and forest affinity info
species_info <- read.csv("Data/species_info.csv")

# # read in NMDS results
# NMDSreslist <- readRDS("NMDSreslist.rds")
```

# NMDS

```{r NMDS}
# list to save plots in
NMDS_plots <- list()

# list to save results in
NMDSreslist <- list() # will be chill to have this saved instead of having to rerun nmds all the time

for(m in names(dissimlist)){
  df <- dissimlist[[m]]
  set.seed(2983847)
  NMDS_res <- metaMDS(df, k = 2, trymax = 999, trace = FALSE) 
  # autotransform by default TRUE
  
  p <- myNMDS(df, NMDS_res, codes_info) + 
    # temporary for figure
    labs(x = "NMDS1", y = "NMDS2")
  
  NMDS_plots[[m]] <- p
  
  NMDSreslist[[m]] <- NMDS_res

}

saveRDS(NMDSreslist, "NMDSreslist.rds")

ggsave("Figures/NMDSplot.svg", NMDS_plots[["df_sitesum"]], width = 16, height = 8) # this one would need a higher concavity of the hulls (cf. functions.R)
ggsave("Figures/NMDSplotFOFF.svg", NMDS_plots[["df_sitesum_FOFF"]], width = 8, height = 8) # 11 food forest and 11 forest sites
ggsave("Figures/NMDSplotFOFF_5.svg", NMDS_plots[["df_sitesum_FOFF_5"]], width = 8, height = 8) # 8 food forest and 8 forest sites
```

```{r NMDS_loc}
set.seed(2983847)
  NMDS_res <- NMDSreslist[["df_sitesum_FOFF_5"]] 
  df <- dissimlist[["df_sitesum_FOFF_5"]]
  # autotransform by default TRUE
  
  p <- myNMDS_loc(df, NMDS_res, codes_info) + 
    # temporary for figure
    labs(x = "NMDS1", y = "NMDS2")
  
ggsave("Figures/NMDSplotFOFF_5_loc.svg", p, width = 8, height = 8) # 8 food forest and 8 forest sites
```

# Indicator species

```{r indicspecies}

for(m in names(specablist)){
  set.seed(2983847)
  df <- (specablist[[m]])
  
  site_class <- data.frame(code = rownames(df)) %>%
    left_join(codes_info) %>%
    select(landuse)

indval <- multipatt(df, site_class$landuse, 
                    control = how(nperm=999)) 

#summary(indval) # for all three cases only Phalangium opilio associated with combination of grass- and cropland land use types
summary(indval, alpha = 0.1) # alpha = 1 to display all
}


```

# Taxonomy

Stacked bars with proportion of individuals belonging to certain taxonomic group.
This is in any case at the site level, not trap level, and taking the mean proportion per site or the proportion after having summed two traps per site gives exactly the same results (logically, but also tested it to be sure).

```{r stackedbars_layout}
# colours
pal <- c("#0C7BDC", "#FFC20A","#40B0A6","#5D3A9B","#FEFE62","#DC3220","#1AFF1A","#D35FB7","#994F00","#88CCEE")

# theme
set_theme <- list(
  theme_bw(),
  theme(panel.grid = element_blank(),
        strip.background.x = element_blank(), 
        strip.placement = "outside",
        axis.line = element_line(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.length.x = unit(0, "pt"),
        panel.border = element_blank(),
        panel.spacing.x = unit(7, "pt"), # does what it implies, space the panels of the facet 
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        strip.text.x = element_text(size = 14))#,
        # legend.text = element_text(size = 13), 
        # legend.title = element_text(size = 14))
)
```

```{r prepare_df}

df_all <- dflist[["df_sitesum"]] %>%
  # filter out all spec. entries except for Leiobunum_spec.
  filter(!(species %in% c("Phalangiidae_spec.",
                          "Nemastomatidae_spec.", 
                          "Nemastoma_spec."))) %>%
  filter(count > 0) %>%
  left_join(codes_info) %>%
  ungroup() %>%
  mutate(landuse = factor(landuse, 
                          levels = c("Food_forest", "Forest",
                                    "Grassland",
                                     "Arable_field" ))) %>%
  # add taxonomic and forest affinity information
  left_join(species_info) %>%
  # add column total count
  group_by(code) %>%
  mutate(tot_count = sum(count)) %>% 
  ungroup()

df_5 <- df_all %>%
  # filter out too small assemblages
  filter(tot_count >= 5/4)
  
# to put info of total count on bar
df_labs <- harvestmen_raw %>% 
    filter(!(species %in% c("Phalangiidae_spec.",
                          "Nemastomatidae_spec.", 
                          "Nemastoma_spec."))) %>%
  group_by(site) %>%
  summarise(tot_count = sum(abundance))

labeller_sites <- setNames(as.character(df_labs$tot_count), df_labs$site)

my_labeller <- labeller(landuse = c("Food_forest" = "Food forest",
                                             "Forest" = "Forest",
                                             "Grassland" = "Grassland",
                                             "Arable_field" = "Cropland"))
```

```{r taxonomy_bars}
# for all sites
tax_stack <- df_all %>%
  ggnested(aes(x = code, y = count, 
               main_group = family, sub_group = genus), #species 
         main_palette = pal) +
  geom_col(position = position_fill(reverse = TRUE)) + 
  # 0 at level 0 of y axis
  scale_y_continuous(expand = expansion(mult = c(0, .05))) + #, labels = scales::percent
  # add padding around outer bars
  scale_x_discrete(expand = expansion(add = .8)) +
  labs(y = "Proportion of captured individuals",
       fill = "Genus") +
  facet_grid(.~ landuse, switch = "x", space = "free_x", scales = "free_x",
             labeller = my_labeller) +
  guides(fill = guide_legend(ncol = 1)) + # number of columns for legend
  set_theme +
  theme(legend.position = "right",
        axis.text.x = element_text(size = 8, angle = 90)) +
  # add padding around outer bars
  scale_x_discrete(expand = expansion(add = .8), 
                   labels = function(x)
      labeller_sites[str_extract(x, regex("[:alpha:]*"))])

ggsave("Figures/tax_stack_all.svg", dpi = 300, unit = "mm", width = 210, height = 160)

# for only those communities with at least 5 individuals (after removal of _spec. entries)
# tax_stack_5 <- df_5 %>%
#   ggnested(aes(x = code, y = count, 
#                main_group = family, sub_group = genus), #species 
#          main_palette = pal) +
#   geom_col(position = position_fill(reverse = TRUE)) + 
#   # 0 at level 0 of y axis
#   scale_y_continuous(expand = expansion(mult = c(0, .05))) + #, labels = scales::percent
#   # add padding around outer bars
#   scale_x_discrete(expand = expansion(add = .8)) +
#   labs(y = "Proportion of captured individuals",
#        fill = "Genus") +
#   facet_grid(.~ landuse, switch = "x", space = "free_x", scales = "free_x",
#              labeller = labeller(landuse = c("Food_forest" = "Food forest",
#                                              "Forest" = "Forest",
#                                              "Grassland" = "Grassland",
#                                              "Arable_field" = "Cropland"))) +
#   guides(fill = guide_legend(ncol = 1)) + # number of columns for legend
#   set_theme +
#   theme(legend.position = "right")
# 
# ggsave("tax_stack_5.png")
```
```{r FO_bar_all}
# for all sites
forest_stack <- dflist[["df_sitesum"]] %>%
  # filter out all spec. entries except for Leiobunum_spec.
  filter(!(species %in% c("Phalangiidae_spec.",
                          "Nemastomatidae_spec.", 
                          "Nemastoma_spec."))) %>%
  filter(count > 0) %>%
  left_join(codes_info) %>%
  ungroup() %>%
  mutate(landuse = factor(landuse, levels = c("Food_forest", "Forest",
                                              "Grassland", "Arable_field" ))) %>%
  # add FO affinity information
  left_join(species_info) %>%
  mutate(forest_affinity = factor(forest_affinity,
                                  levels = c("Strong-closed", "Strong",
                                             "Strong-open","Moderate",
                                             "Weak","None", "NA"))) %>%
  ggplot(aes(x = code, y = count, fill = forest_affinity)) +
  geom_col(position = position_fill(reverse = TRUE)) + 
  # 0 at level 0 of y axis
  scale_y_continuous(expand = expansion(mult = c(0, .05))) +
  # add padding around outer bars
  scale_x_discrete(expand = expansion(add = .8)) +
  scale_fill_manual(values = c("#275317","#3B7D23","#8ED973",
                               "#0B76A0","#96DCF8","#D9D9D9",
                               "#ededed")) +
  labs(y = "Proportion of captured individuals",
       fill = "Forest affinity") +
  # 0 at level 0 of y axis
  scale_y_continuous(expand = expansion(mult = c(0, .05))) + #, labels = scales::percent
  # add padding around outer bars
  scale_x_discrete(expand = expansion(add = .8)) +
  facet_grid(.~ landuse, switch = "x", space = "free_x", scales = "free_x",
             labeller = labeller(landuse = c("Food_forest" = "Food forest",
                                             "Forest" = "Forest",
                                             "Grassland" = "Grassland",
                                             "Arable_field" = "Cropland"))) +
  guides(fill = guide_legend(ncol = 1)) + # number of columns for legend
  set_theme +
  theme(legend.position = "right")

ggsave("Figures/FO_stack_all.svg", 
         dpi = 300, unit = "mm", width = 210, height = 160)
```

```{r habitat_bar_all}
# for all sites
habitat_stack <- dflist[["df_sitesum"]] %>%
  # filter out all spec. entries except for Leiobunum_spec.
  filter(!(species %in% c("Phalangiidae_spec.",
                          "Nemastomatidae_spec.", 
                          "Nemastoma_spec."))) %>%
  filter(count > 0) %>%
  left_join(codes_info) %>%
  ungroup() %>%
  mutate(landuse = factor(landuse, levels = c("Food_forest", "Forest",
                                              "Grassland", "Arable_field" ))) %>%
  # add FO affinity information
  left_join(species_info) %>%
  mutate(habitat_affinity = factor(habitat_affinity,
                                  levels = c("Forest-closed", "Forest",
                                             "Forest-open","Mixed-equal",
                                             "Mixed-open","Open", NA))) %>%
  ggplot(aes(x = code, y = count, fill = habitat_affinity)) +
  geom_col(position = position_fill(reverse = TRUE)) + 
  # 0 at level 0 of y axis
  scale_y_continuous(expand = expansion(mult = c(0, .05))) +
  # add padding around outer bars
  scale_x_discrete(expand = expansion(add = .8)) +
  scale_fill_manual(values = c("#275317","#3B7D23","#6a9731",
                               "#98b142","#c8cb57","#fae470",
                               "#D9D9D9"),
                    na.value = "#D9D9D9") +
  labs(y = "Proportion of captured individuals",
       fill = "Preferred habitat") +
  # 0 at level 0 of y axis
  scale_y_continuous(expand = expansion(mult = c(0, .05))) + #, labels = scales::percent
  # add padding around outer bars
  scale_x_discrete(expand = expansion(add = .8)) +
  facet_grid(.~ landuse, switch = "x", space = "free_x", scales = "free_x",
             labeller = labeller(landuse = c("Food_forest" = "Food forest",
                                             "Forest" = "Forest",
                                             "Grassland" = "Grassland",
                                             "Arable_field" = "Cropland"))) +
  guides(fill = guide_legend(ncol = 1)) + # number of columns for legend
  set_theme +
  theme(legend.position = "right",
        axis.text.x = element_text(size = 8, angle = 90)) +
  # add padding around outer bars
  scale_x_discrete(expand = expansion(add = .8), 
                   labels = function(x)
      labeller_sites[str_extract(x, regex("[:alpha:]*"))])

ggsave("Figures/habitat_stack_all.svg", 
         dpi = 300, unit = "mm", width = 210, height = 160)
```


# Species table

```{r speciestable}

# averaged across sessions (still need to fix this, sum per species, now multiple rows per species)
# tbl <- dflist[["df_sitesum"]] %>% 
#   # here count is the average of the four or three sessions
#   # could also opt for displaying total number of individuals, not averaged?
#   # but this is also because perhaps bias in land use types in terms of visitation session 4?
#   left_join(codes_info) %>%
#   filter(count > 0) %>%
#   mutate(landuse = factor(landuse, levels = c("Food_forest", "Forest", 
#                                               "Grassland", "Arable_field" ))) %>%
#   ungroup() %>%
#   pivot_wider(names_from = landuse,
#               values_from = count) %>%
#   replace_na(list(Food_forest = 0, Forest = 0, Grassland = 0, Arable_field = 0)) %>%
#   mutate(Total = rowSums(across(where(is.numeric)))) %>%
#   # round
#   left_join(species_info) %>%
#   arrange(family) %>%
#   select(species, family, 
#          forest_affinity, 
#          #IUCN_BE, IUCN_NL, # may still look for this, but likely doesn't exist
#          Food_forest, Forest, Grassland, Arable_field, Total) %>%
#   # cleaner presentation of strings
#   mutate(species = gsub("_", " ", species)) %>%
#   rename("Species" = "species",
#          "Affinity to forests" = "forest_affinity",
#          "Food forest" = "Food_forest",
#          "Cropland" = "Arable_field",
#          )


# not averaged across sessions 
tbl_total <- readRDS("harvestmen.rds") %>%
  left_join(codes_info %>% select(site, location, landuse) %>% distinct()) %>%
  # sum all counts per species per land use type
  group_by(landuse, species) %>%
  summarise(count = sum(abundance)) %>%
  mutate(landuse = factor(landuse, levels = c("Food_forest", "Forest", 
                                              "Grassland", "Arable_field" ))) %>%
  ungroup() %>%
  pivot_wider(names_from = landuse,
              values_from = count) %>%
  replace_na(list(Food_forest = 0, Forest = 0, Grassland = 0, Arable_field = 0)) %>%
  mutate(Total = rowSums(across(where(is.numeric)))) %>%
  left_join(species_info) %>%
  arrange(family) %>%
  select(species, family, 
         forest_affinity, 
         #IUCN_BE, IUCN_NL, # may still look for this, but likely doesn't exist
         Food_forest, Forest, Grassland, Arable_field, Total) %>%
  # cleaner presentation of strings
  mutate(species = gsub("_", " ", species)) %>%
  rename("Species" = "species",
         "Affinity to forests" = "forest_affinity",
         "Food forest" = "Food_forest",
         "Cropland" = "Arable_field",
         )
  
  
# save as csv
write.csv(tbl_total, "speciestable_total.csv")
```


## PERMDISP and PERMANOVA

### 3 betadispers

```{r PERMDISP}
# for now only doing it for case with
# only food forest and forest sites
# only sites with at least 5 individuals in total (8 FF sites, 8 FO sites)
# not including location, because this doesn't work (several locations with just one site)
betadisper_output <- data.frame()

  set.seed(2983847)
  df <- as.matrix(dissimlist[["df_sitesum_FOFF_5"]])
  rn <- data.frame(code = rownames(df)) %>%
    left_join(codes_info %>% select(code, landuse, location))
  
  n_obs <- rn %>%
    group_by(landuse) %>%
    summarise(n = n())
  
    dissim <- rn %>%
        filter(landuse %in% c("Food_forest", "Forest")) %>%
        pull(code) %>%
      { df[.,.] } %>%
      as.dist()
    dissim_info <- rn %>% filter(landuse %in% c("Food_forest", "Forest"))
    disper_lu <- betadisper(dissim, dissim_info$landuse, type="centroid")
    FF <- round(disper_lu$group.distances[["Food_forest"]], digits=2)
    CON <- round(disper_lu$group.distances[["Forest"]], digits=2)
    F_lu <- format(round(permutest(disper_lu)$tab$`F`[1], 
                         digits = 1))
    p_lu <- permutest(disper_lu)$tab$`Pr(>F)`[1]

    output <- cbind( 
                    contrast = "Forest",
                    n_FF = n_obs %>% filter(landuse == "Food_forest") %>% pull(),
                    n_CON = n_obs %>% filter(landuse == "Forest") %>% pull(),
                     FF, CON, F_lu, p_lu
                    )
    betadisper_output <- rbind(betadisper_output, output)


flextable(betadisper_output)
write.csv(betadisper_output, "PERMDISP_output.csv")
```

### 3 permanovas

```{r PERMANOVA}
permanova_output <- data.frame()

  set.seed(2983847)
  df <- as.matrix(dissimlist[["df_sitesum_FOFF_5"]])
  rn <- data.frame(code = rownames(as.matrix(df))) %>%
    left_join(codes_info %>% select(code, landuse, location))
  
  n_obs <- rn %>%
    group_by(landuse) %>%
    summarise(n = n()) 

  for(l in setdiff(n_obs$landuse, "Food_forest")){ # only for those reference land use types with >= 6 obs.
      dissim <- rn %>%
          filter(landuse %in% c("Food_forest",l)) %>%
          pull(code) %>%
            { df[.,.] }
      dissim_info <- rn %>% filter(landuse=="Food_forest"|landuse==l)
      
      results <- adonis2(dissim ~ landuse, 
                          data = dissim_info, by="terms")
      
      results_lu <- results[1,] %>%
        rename("p" = "Pr(>F)",
               "pseudo_F" = "F") %>%
        mutate(df = paste0(Df, ", ", results[3,1])) %>%
        select(df, pseudo_F, p, R2)
      
      
      permanova_output <- rbind(permanova_output, round_df(results_lu,digits=3))

    }
  
flextable(permanova_output)
write.csv(permanova_output, "PERMANOVA_output.csv")
```
-->
