---
title: "DA_quant_div"
author: "Lieke"
date: "2026-03-16"
output: html_document
---

```{r packages}
library(tidyverse)
# modelling
library(glmmTMB)
# diagnostics
library(DHARMa)
library(performance)
library(see)
# for bootstrapping
# run on parallel cores
library(future)
library(future.apply)
# for bootMer
library(lme4)
# plotting
library(patchwork)
# reporting
library(report)
library(emmeans)
```

```{r functions}
source("functions.R")
boot_pval <- function(x) { 2*min(mean(x<0), mean(x>0)) } 
```

For the comparison of the activity-density and diversity of harvestmen between food forests and the reference land use types, we use the data at the site level. This means that for each session, the two traps for one site are summed, and that the species counts are then averaged across the four sessions (to account for sites not being sampled in specific sessions, make fair comparison).

For the activity-density, we use all individuals, including the juveniles. For the analysis of the diversity (as for the community composition), we exclude the _spec. entries.

```{r data}
codes_info <- read.csv("Data/codes_info.csv")

quant <- readRDS("quantitylist.rds")[["df_sitesum"]] %>%
  left_join(codes_info) %>%
  mutate(landuse = factor(landuse,
                          levels = c("Food_forest", "Forest",
                                     "Grassland", "Arable_field")))
         # manually checked value for site AK and correct (27/4 = 6.75)

div <- readRDS("diversitylist.rds")[["df_sitesum"]] %>%
  left_join(codes_info) %>%
  mutate(landuse = factor(landuse,
                          levels = c("Food_forest", "Forest",
                                     "Grassland", "Arable_field")))
# manually checked values for site AK and correct (in total, across 4 sessions, 3 species (non _spec. entries))

bb_q <- readRDS("bb_q.rds")
bb_D0 <- readRDS("bb_D0.rds")
bb_D1 <- readRDS("bb_D1.rds")
```

```{r output_lists}
# list for plots
plots <- list()

# df for ratios and their 95%CI
ratios <- data.frame()

# df for model summary
modeloutputs <- data.frame()
```

# Quantity

Data contains zeroes, but is otherwise continuous positive (continuous because averaged) --> tweedie distribution.

```{r q_model_diagn}
# Tweedie with RE
model <- glmmTMB(count ~ landuse + (1|location),
                  family = tweedie,
                  data = quant)
plot(simulateResiduals(model)) # looks OK!
r2(model) 
```

```{r q_model_bootstrap}
plan(multisession, workers = parallel::detectCores() - 1)

# parametric bootstrapping
bb_q <- bootMer(model,
              FUN = function(x) {
                ff <- fixef(x)[["cond"]] 
                ff 
              },
              nsim = 5000,
              parallel = "future",
              ncpus = parallel::detectCores() - 1)

# check - look kind of OK
bb_t <- as.data.frame(bb_q$t) %>%
    pivot_longer(cols = everything())
  
for(n in unique(bb_t$name)){
  # overall histograms
  print(bb_t %>% 
          filter(name == n) %>%
          ggplot(aes(x = value)) +
            geom_histogram())
}
      
# save
saveRDS(bb_q, "bb_q.rds")
```

```{r q_model_output}
## "extra dispersion" attributable to random effects 
## (for bias-adjustment when back-transforming)
  extra_disp <- insight::get_variance(model)[["var.intercept"]]
  
  # put model summary into dataframe
  reptable <- as.data.frame(report_table(model)) # this gives the Nakagawa-R²s, as you would obtain them using performance::r2()
  output <- data.frame(variable = "activity-density",
                          RE_variance = reptable %>%
                        filter(Effects == "random") %>%
                        pull(Coefficient),
                      R2marg = reptable %>%
                        filter(Parameter == "R2 (marginal)") %>%
                        pull(Fit),
                      R2cond = reptable %>%
                        filter(Parameter == "R2 (conditional)") %>%
                        pull(Fit),
                      sigma = reptable %>%
                        filter(Parameter == "Sigma") %>% # is the same as the dispersion parameter
                        pull(Fit))
  
  modeloutputs <- rbind(modeloutputs, output)
  
    
  # process bootstrap values 
  # to obtain 95%CIs of means and ratios in response scale and p-values
  ## make dataframe from bootstrap estimates, with landuse means and differences in log scale
  bb_df_log <- as.data.frame(bb_q[["t"]]) %>%
    # rename columns
    rename("Food_forest_mean" = "(Intercept)") %>%
    rename_with(.fn = ~ gsub("landuse", "diff_", .x), starts_with("landuse")) %>% 
    # add column with means of reference landuses, 
    # by adding their estimated difference to the estimated mean of the food forest
    mutate(Forest_mean = Food_forest_mean + diff_Forest,
           Grassland_mean = Food_forest_mean + diff_Grassland,
           Arable_field_mean = Food_forest_mean + diff_Arable_field,
           .after = "Food_forest_mean")

  ## make dataframe with landuse means in response scale, and take FF/ref ratio from those 
  bb_df_response <- bb_df_log %>%
    select(ends_with("mean")) %>%
    # back-transform with bias adjustment for RE variance
    mutate(across(everything(), ~ exp(.) + 0.5*exp(.)*extra_disp^2)) %>%
    # compute ratio from these bootstrap values for the estimated means in the response scale
    mutate(ratio_Forest = Food_forest_mean/Forest_mean,
           ratio_Grassland = Food_forest_mean/Grassland_mean,
           ratio_Arable_field = Food_forest_mean/Arable_field_mean)
  
  ## obtain CIs for model estimated means for plotting
  ## we plot the model estimated mean and make the CI around this by using the SD obtained from the parametric bootstrapping 
  
  mean_boot_sd <- bb_df_response %>%
    select(ends_with("mean")) %>%
    pivot_longer(names_to = "landuse", names_pattern = "(.*)_mean",
                 cols = ends_with("mean")) %>% 
    group_by(landuse) %>%
    reframe(boot_sd = sd(value))
  
  estimates_response <- data.frame(model_estimate = fixef(model)[["cond"]]) %>% 
    rownames_to_column(var = "landuse") %>%
    # change landuse names
    mutate(landuse = case_when(landuse == "(Intercept)" ~ "Food_forest",
                             .default = str_remove(landuse, "^landuse"))) %>%
    # compute actual means for reference land uses (i.e. subtract from food forest mean)
    mutate(model_mean = ifelse(landuse != "Food_forest", 
                             model_estimate[landuse == "Food_forest"] + model_estimate,
                             model_estimate)) %>%
    # back-transform to response scale, with bias adjustment
    mutate(across(where(is.numeric), ~ exp(.) + 0.5*exp(.)*extra_disp^2)) %>%
    # join with sd values
    left_join(mean_boot_sd) %>%
    # add CI limits
    mutate(lower.CL = model_mean - 1.96*boot_sd,
           upper.CL = model_mean + 1.96*boot_sd) %>%
    # format for ggplot_the_response
    select(landuse, model_mean, lower.CL, upper.CL) %>%
    rename(response = model_mean)
    
  
  ## obtain CIs for ratios - both by using sd from parametric bootstrapping and by using Wald CIs
  ratio_boot_sd <- bb_df_response %>%
    select(starts_with("ratio")) %>%
    pivot_longer(cols = starts_with("ratio"),
                 names_to = "landuse", 
                 names_pattern = "ratio_(.*)")%>%
    group_by(landuse) %>%
    reframe(boot_sd = sd(value))
  
  ratio_Wald_interval <- confint(contrast(emmeans(model, ~landuse), 
                                          "trt.vs.ctrl", reverse = TRUE, type = "response"))
  
  CI_ratio_response <- estimates_response %>%
    select(landuse, response) %>%
    mutate(ratio = response[landuse == "Food_forest"]/response) %>%
    left_join(ratio_boot_sd) %>%
    filter(!landuse == "Food_forest") %>%
    mutate(lower.CL_boot = ratio - 1.96*boot_sd,
           upper.CL_boot = ratio + 1.96*boot_sd) %>%
    # add CI limits from asymptotic Wald intervals
    mutate(lower.CL_Wald = ratio_Wald_interval[,"asymp.LCL"],
           upper.CL_Wald = ratio_Wald_interval[,"asymp.UCL"])
    
  ## obtain two-tailed p-values, 
  boot_contrasts <- data.frame(p_raw = apply(bb_q$t[, colnames(bb_q$t) != "(Intercept)"], 
                                             2, boot_pval)) %>%
    # adjust p-values for multiple comparisons using Holm's method (FWER) 
    mutate(p_adj_Holm = p.adjust(p_raw, method = "holm"))  %>% 
    # already format to supply to ggplot_the_response
    # adding column contrast
    rownames_to_column(var = "contrast") %>%
    mutate(contrast = paste0(str_remove(contrast, "^landuse"), " - Food_forest")) %>%
    rename(p.value = p_adj_Holm)

  ## plotting
  p <- ggplot_the_response(
    fit = model,
    fit_emm = estimates_response, 
    fit_pairs = boot_contrasts, 
        dots = "sina1",
      contrast_rows = "all",
      palette = pal_okabe_ito_FF,
      y_label = "Mean activity-density per session",
      g_label = "none"
    ) + 
      theme(axis.title.x = element_blank(),
            axis.title.y = element_text(size = 16),
            axis.text = element_text(size = 14) 
            ) +     
      scale_x_discrete(labels = c("Food forest", "Forest", 
                                  "Grassland", "Cropland"))
  
  # add dashed horizontal line at y = 0, in a way that puts line below other layers
  p$layers <- c(geom_hline(yintercept = 0, linetype = "dashed", colour = "darkgrey"), p$layers)
  
  # add plot to list
  plots[["quantity"]] <- p
    
  # format ratios for table output
  CI_ratio_response <- CI_ratio_response %>%
    mutate(variable = "activity-density",
           ratio_CI_boot = paste0(round(ratio, digits = 2),
                               " [", 
                               round(lower.CL_boot, digits = 2),
                               ", ", 
                               round(upper.CL_boot, digits = 2),
                               "]"),
           ratio_CI_Wald = paste0(round(ratio, digits = 2),
                               " [", 
                               round(lower.CL_Wald, digits = 2),
                               ", ", 
                               round(upper.CL_Wald, digits = 2),
                               "]")) %>%
    select(variable, landuse, ratio_CI_boot, ratio_CI_Wald) %>%
    pivot_longer(cols = c(ratio_CI_boot, ratio_CI_Wald),
                 values_to = "ratio_CI",
                 names_to = "method",
                 names_pattern = ".*_(.*)") %>%
      pivot_wider(names_from = landuse,
                values_from = ratio_CI) %>%
    rename_with(.cols = -method,
                .fn = ~paste0("Food forest/", .))
  
  # put in dataframe
  ratios <- rbind(ratios, CI_ratio_response)
```

# Number of species

Count variable --> Poisson distribution or negative binomial in case of overdispersion

```{r D0_model_diagn}
# Poisson with RE
model <- glmmTMB(D0 ~ landuse + (1|location),
                  family = poisson,
                  data = div)
plot(simulateResiduals(model)) # looks good!
r2(model) 
```

```{r D0_model_bootstrap}
plan(multisession, workers = parallel::detectCores() - 1)

# parametric bootstrapping
bb_D0 <- bootMer(model,
              FUN = function(x) {
                ff <- fixef(x)[["cond"]] 
                ff 
              },
              nsim = 5000,
              parallel = "future",
              ncpus = parallel::detectCores() - 1)

# check
bb_t <- as.data.frame(bb_D0$t) %>%
    pivot_longer(cols = everything())
  
    for(n in unique(bb_t$name)){
      # overall histograms
      print(bb_t %>% 
              filter(name == n) %>%
              ggplot(aes(x = value)) +
                geom_histogram())
    }

# save
saveRDS(bb_D0, "bb_D0.rds")
```

```{r D0_model_output}
## "extra dispersion" attributable to random effects 
## (for bias-adjustment when back-transforming)
  extra_disp <- insight::get_variance(model)[["var.intercept"]]
  
  # put model summary into dataframe
  reptable <- as.data.frame(report_table(model)) # this gives the Nakagawa-R²s, as you would obtain them using performance::r2()
  output <- data.frame(variable = "0D",
                          RE_variance = reptable %>%
                        filter(Effects == "random") %>%
                        pull(Coefficient),
                      R2marg = reptable %>%
                        filter(Parameter == "R2 (marginal)") %>%
                        pull(Fit),
                      R2cond = reptable %>%
                        filter(Parameter == "R2 (conditional)") %>%
                        pull(Fit),
                      sigma = reptable %>%
                        filter(Parameter == "Sigma") %>% # is the same as the dispersion parameter
                        pull(Fit))
  
  modeloutputs <- rbind(modeloutputs, output)
  
    
  # process bootstrap values 
  # to obtain 95%CIs of means and ratios in response scale and p-values
  ## make dataframe from bootstrap estimates, with landuse means and differences in log scale
  bb_df_log <- as.data.frame(bb_D0[["t"]]) %>%
    # rename columns
    rename("Food_forest_mean" = "(Intercept)") %>%
    rename_with(.fn = ~ gsub("landuse", "diff_", .x), starts_with("landuse")) %>% 
    # add column with means of reference landuses, 
    # by adding their estimated difference to the estimated mean of the food forest
    mutate(Forest_mean = Food_forest_mean + diff_Forest,
           Grassland_mean = Food_forest_mean + diff_Grassland,
           Arable_field_mean = Food_forest_mean + diff_Arable_field,
           .after = "Food_forest_mean")

  ## make dataframe with landuse means in response scale, and take FF/ref ratio from those 
  bb_df_response <- bb_df_log %>%
    select(ends_with("mean")) %>%
    # back-transform with bias adjustment for RE variance
    mutate(across(everything(), ~ exp(.) + 0.5*exp(.)*extra_disp^2)) %>%
    # compute ratio from these bootstrap values for the estimated means in the response scale
    mutate(ratio_Forest = Food_forest_mean/Forest_mean,
           ratio_Grassland = Food_forest_mean/Grassland_mean,
           ratio_Arable_field = Food_forest_mean/Arable_field_mean)
  
  ## obtain CIs for model estimated means for plotting
  ## we plot the model estimated mean and make the CI around this by using the SD obtained from the parametric bootstrapping 
  
  mean_boot_sd <- bb_df_response %>%
    select(ends_with("mean")) %>%
    pivot_longer(names_to = "landuse", names_pattern = "(.*)_mean",
                 cols = ends_with("mean")) %>% 
    group_by(landuse) %>%
    reframe(boot_sd = sd(value))
  
  estimates_response <- data.frame(model_estimate = fixef(model)[["cond"]]) %>% 
    rownames_to_column(var = "landuse") %>%
    # change landuse names
    mutate(landuse = case_when(landuse == "(Intercept)" ~ "Food_forest",
                             .default = str_remove(landuse, "^landuse"))) %>%
    # compute actual means for reference land uses (i.e. subtract from food forest mean)
    mutate(model_mean = ifelse(landuse != "Food_forest", 
                             model_estimate[landuse == "Food_forest"] + model_estimate,
                             model_estimate)) %>%
    # back-transform to response scale, with bias adjustment
    mutate(across(where(is.numeric), ~ exp(.) + 0.5*exp(.)*extra_disp^2)) %>%
    # join with sd values
    left_join(mean_boot_sd) %>%
    # add CI limits
    mutate(lower.CL = model_mean - 1.96*boot_sd,
           upper.CL = model_mean + 1.96*boot_sd) %>%
    # format for ggplot_the_response
    select(landuse, model_mean, lower.CL, upper.CL) %>%
    rename(response = model_mean)
    
  
  ## obtain CIs for ratios - both by using sd from parametric bootstrapping and by using Wald CIs
  ratio_boot_sd <- bb_df_response %>%
    select(starts_with("ratio")) %>%
    pivot_longer(cols = starts_with("ratio"),
                 names_to = "landuse", 
                 names_pattern = "ratio_(.*)")%>%
    group_by(landuse) %>%
    reframe(boot_sd = sd(value))
  
  ratio_Wald_interval <- confint(contrast(emmeans(model, ~landuse), 
                                          "trt.vs.ctrl", reverse = TRUE, type = "response"))
  
  CI_ratio_response <- estimates_response %>%
    select(landuse, response) %>%
    mutate(ratio = response[landuse == "Food_forest"]/response) %>%
    left_join(ratio_boot_sd) %>%
    filter(!landuse == "Food_forest") %>%
    mutate(lower.CL_boot = ratio - 1.96*boot_sd,
           upper.CL_boot = ratio + 1.96*boot_sd) %>%
    # add CI limits from asymptotic Wald intervals
    mutate(lower.CL_Wald = ratio_Wald_interval[,"asymp.LCL"],
           upper.CL_Wald = ratio_Wald_interval[,"asymp.UCL"])
    
  ## obtain two-tailed p-values, 
  boot_contrasts <- data.frame(p_raw = apply(bb_D0$t[, colnames(bb_D0$t) != "(Intercept)"], 
                                             2, boot_pval)) %>%
    # adjust p-values for multiple comparisons using Holm's method (FWER) 
    mutate(p_adj_Holm = p.adjust(p_raw, method = "holm"))  %>% 
    # already format to supply to ggplot_the_response
    # adding column contrast
    rownames_to_column(var = "contrast") %>%
    mutate(contrast = paste0(str_remove(contrast, "^landuse"), " - Food_forest")) %>%
    rename(p.value = p_adj_Holm)

  ## plotting
  p <- ggplot_the_response(
    fit = model,
    fit_emm = estimates_response, 
    fit_pairs = boot_contrasts, 
        dots = "sina1",
      contrast_rows = "all",
      palette = pal_okabe_ito_FF,
      y_label = expression("Number of species ("^0*"D)"),
      g_label = "none"
    ) + 
      theme(axis.title.x = element_blank(),
            axis.title.y = element_text(size = 16),
            axis.text = element_text(size = 14) 
            ) +     
      scale_x_discrete(labels = c("Food forest", "Forest", 
                                  "Grassland", "Cropland"))
  
  # add dashed horizontal line at y = 0, in a way that puts line below other layers
  p$layers <- c(geom_hline(yintercept = 0, linetype = "dashed", colour = "darkgrey"), p$layers)
  
  # add plot to list
  plots[["0D"]] <- p
    
  # format ratios for table output
  CI_ratio_response <- CI_ratio_response %>%
    mutate(variable = "0D",
           ratio_CI_boot = paste0(round(ratio, digits = 2),
                               " [", 
                               round(lower.CL_boot, digits = 2),
                               ", ", 
                               round(upper.CL_boot, digits = 2),
                               "]"),
           ratio_CI_Wald = paste0(round(ratio, digits = 2),
                               " [", 
                               round(lower.CL_Wald, digits = 2),
                               ", ", 
                               round(upper.CL_Wald, digits = 2),
                               "]")) %>%
    select(variable, landuse, ratio_CI_boot, ratio_CI_Wald) %>%
    pivot_longer(cols = c(ratio_CI_boot, ratio_CI_Wald),
                 values_to = "ratio_CI",
                 names_to = "method",
                 names_pattern = ".*_(.*)") %>%
      pivot_wider(names_from = landuse,
                values_from = ratio_CI) %>%
    rename_with(.cols = -method,
                .fn = ~paste0("Food forest/", .))
  
  # put in dataframe
  ratios <- rbind(ratios, CI_ratio_response)

```

# Effective number of species

Data contains zeroes, but is otherwise continuous positive --> tweedie distribution.

```{r D1_model_diagn}
# Tweedie without RE (RE variance too small) 
model <- glmmTMB(D1 ~ landuse,
                  family = tweedie,
                  data = div)
plot(simulateResiduals(model)) # looks OK!

library(rcompanion)
nagelkerke(model)
```

```{r D1_model_bootstrap}
plan(multisession, workers = parallel::detectCores() - 1)

# parametric bootstrapping
bb_D1 <- bootMer(model,
              FUN = function(x) {
                ff <- fixef(x)[["cond"]] 
                ff 
              },
              nsim = 5000,
              parallel = "future",
              ncpus = parallel::detectCores() - 1)

# check
bb_t <- as.data.frame(bb_D1$t) %>%
    pivot_longer(cols = everything())
  
    for(n in unique(bb_t$name)){
      # overall histograms
      print(bb_t %>% 
              filter(name == n) %>%
              ggplot(aes(x = value)) +
                geom_histogram())
    }

# save
saveRDS(bb_D1, "bb_D1.rds")
```

```{r D1_model_output}
# ## "extra dispersion" attributable to random effects 
# ## (for bias-adjustment when back-transforming)
#   extra_disp <- insight::get_variance(model)[["var.intercept"]]
  
  # put model summary into dataframe
  # reptable <- as.data.frame(report_table(model)) # this gives the Nakagawa-R²s, as you would obtain them using performance::r2()
  # output <- data.frame(variable = "1D",
  #                         RE_variance = reptable %>%
  #                       filter(Effects == "random") %>%
  #                       pull(Coefficient),
  #                     R2marg = reptable %>%
  #                       filter(Parameter == "R2 (marginal)") %>%
  #                       pull(Fit),
  #                     R2cond = reptable %>%
  #                       filter(Parameter == "R2 (conditional)") %>%
  #                       pull(Fit),
  #                     sigma = reptable %>%
  #                       filter(Parameter == "Sigma") %>% # is the same as the dispersion parameter
  #                       pull(Fit))
  
  #modeloutputs <- rbind(modeloutputs, output)
  
    
  # process bootstrap values 
  # to obtain 95%CIs of means and ratios in response scale and p-values
  ## make dataframe from bootstrap estimates, with landuse means and differences in log scale
  bb_df_log <- as.data.frame(bb_D1[["t"]]) %>%
    # rename columns
    rename("Food_forest_mean" = "(Intercept)") %>%
    rename_with(.fn = ~ gsub("landuse", "diff_", .x), starts_with("landuse")) %>% 
    # add column with means of reference landuses, 
    # by adding their estimated difference to the estimated mean of the food forest
    mutate(Forest_mean = Food_forest_mean + diff_Forest,
           Grassland_mean = Food_forest_mean + diff_Grassland,
           Arable_field_mean = Food_forest_mean + diff_Arable_field,
           .after = "Food_forest_mean")

  ## make dataframe with landuse means in response scale, and take FF/ref ratio from those 
  bb_df_response <- bb_df_log %>%
    select(ends_with("mean")) %>%
    # back-transform with bias adjustment for RE variance
    mutate(across(everything(), ~ exp(.) + 0.5*exp(.)*extra_disp^2)) %>%
    # compute ratio from these bootstrap values for the estimated means in the response scale
    mutate(ratio_Forest = Food_forest_mean/Forest_mean,
           ratio_Grassland = Food_forest_mean/Grassland_mean,
           ratio_Arable_field = Food_forest_mean/Arable_field_mean)
  
  ## obtain CIs for model estimated means for plotting
  ## we plot the model estimated mean and make the CI around this by using the SD obtained from the parametric bootstrapping 
  
  mean_boot_sd <- bb_df_response %>%
    select(ends_with("mean")) %>%
    pivot_longer(names_to = "landuse", names_pattern = "(.*)_mean",
                 cols = ends_with("mean")) %>% 
    group_by(landuse) %>%
    reframe(boot_sd = sd(value))
  
  estimates_response <- data.frame(model_estimate = fixef(model)[["cond"]]) %>% 
    rownames_to_column(var = "landuse") %>%
    # change landuse names
    mutate(landuse = case_when(landuse == "(Intercept)" ~ "Food_forest",
                             .default = str_remove(landuse, "^landuse"))) %>%
    # compute actual means for reference land uses (i.e. subtract from food forest mean)
    mutate(model_mean = ifelse(landuse != "Food_forest", 
                             model_estimate[landuse == "Food_forest"] + model_estimate,
                             model_estimate)) %>%
    # back-transform to response scale, with bias adjustment
    mutate(across(where(is.numeric), ~ exp(.) + 0.5*exp(.)*extra_disp^2)) %>%
    # join with sd values
    left_join(mean_boot_sd) %>%
    # add CI limits
    mutate(lower.CL = model_mean - 1.96*boot_sd,
           upper.CL = model_mean + 1.96*boot_sd) %>%
    # format for ggplot_the_response
    select(landuse, model_mean, lower.CL, upper.CL) %>%
    rename(response = model_mean)
    
  
  ## obtain CIs for ratios - both by using sd from parametric bootstrapping and by using Wald CIs
  ratio_boot_sd <- bb_df_response %>%
    select(starts_with("ratio")) %>%
    pivot_longer(cols = starts_with("ratio"),
                 names_to = "landuse", 
                 names_pattern = "ratio_(.*)")%>%
    group_by(landuse) %>%
    reframe(boot_sd = sd(value))
  
  ratio_Wald_interval <- confint(contrast(emmeans(model, ~landuse), 
                                          "trt.vs.ctrl", reverse = TRUE, type = "response"))
  
  CI_ratio_response <- estimates_response %>%
    select(landuse, response) %>%
    mutate(ratio = response[landuse == "Food_forest"]/response) %>%
    left_join(ratio_boot_sd) %>%
    filter(!landuse == "Food_forest") %>%
    mutate(lower.CL_boot = ratio - 1.96*boot_sd,
           upper.CL_boot = ratio + 1.96*boot_sd) %>%
    # add CI limits from asymptotic Wald intervals
    mutate(lower.CL_Wald = ratio_Wald_interval[,"asymp.LCL"],
           upper.CL_Wald = ratio_Wald_interval[,"asymp.UCL"])
    
  ## obtain two-tailed p-values, 
  boot_contrasts <- data.frame(p_raw = apply(bb_D1$t[, colnames(bb_D1$t) != "(Intercept)"], 
                                             2, boot_pval)) %>%
    # adjust p-values for multiple comparisons using Holm's method (FWER) 
    mutate(p_adj_Holm = p.adjust(p_raw, method = "holm"))  %>% 
    # already format to supply to ggplot_the_response
    # adding column contrast
    rownames_to_column(var = "contrast") %>%
    mutate(contrast = paste0(str_remove(contrast, "^landuse"), " - Food_forest")) %>%
    rename(p.value = p_adj_Holm)

  ## plotting
  p <- ggplot_the_response(
    fit = model,
    fit_emm = estimates_response, 
    fit_pairs = boot_contrasts, 
        dots = "sina1",
      contrast_rows = "all",
      palette = pal_okabe_ito_FF,
      y_label = expression("Effective number of species ("^1*"D)"),
      g_label = "none"
    ) + 
      theme(axis.title.x = element_blank(),
            axis.title.y = element_text(size = 16),
            axis.text = element_text(size = 14) 
            ) +     
      scale_x_discrete(labels = c("Food forest", "Forest", 
                                  "Grassland", "Cropland"))
  
  # add dashed horizontal line at y = 0, in a way that puts line below other layers
  p$layers <- c(geom_hline(yintercept = 0, linetype = "dashed", colour = "darkgrey"), p$layers)
  
  # add plot to list
  plots[["1D"]] <- p
    
  # format ratios for table output
  CI_ratio_response <- CI_ratio_response %>%
    mutate(variable = "1D",
           ratio_CI_boot = paste0(round(ratio, digits = 2),
                               " [", 
                               round(lower.CL_boot, digits = 2),
                               ", ", 
                               round(upper.CL_boot, digits = 2),
                               "]"),
           ratio_CI_Wald = paste0(round(ratio, digits = 2),
                               " [", 
                               round(lower.CL_Wald, digits = 2),
                               ", ", 
                               round(upper.CL_Wald, digits = 2),
                               "]")) %>%
    select(variable, landuse, ratio_CI_boot, ratio_CI_Wald) %>%
    pivot_longer(cols = c(ratio_CI_boot, ratio_CI_Wald),
                 values_to = "ratio_CI",
                 names_to = "method",
                 names_pattern = ".*_(.*)") %>%
      pivot_wider(names_from = landuse,
                values_from = ratio_CI) %>%
    rename_with(.cols = -method,
                .fn = ~paste0("Food forest/", .))
  
  # put in dataframe
  ratios <- rbind(ratios, CI_ratio_response)
```

# Saving output

```{r save_output}
# plots
(compound_plot <- wrap_plots(plots[c("quantity", "0D","1D")], ncol = 3))

ggsave("Fig_2.svg", compound_plot, width = 360, height = 160, 
       unit = "mm", dpi = 600)

# table with ratios and their 95%CI
write.csv(ratios, "ratios.csv")

# df for model summary
mutate(modeloutputs, across(where(is.numeric), round, 2)) %>%
  write.csv("modeloutputs.csv")
```

