---
title: "Prep_communitycomposition"
author: "Lieke"
date: "2024-12-08"
output: html_document
---

```{r packages}
# general
library(tidyverse)
# diversity indices
library(iNEXT)
library(hillR)
# dissimilarity matrix, NMDS, PERMANOVA, PERMDISP
library(vegan)
```

```{r functions}
source("functions.R")
```

# Data

```{r additional_data}
codes_info <- read.csv("Data/codes_info.csv")
sites_info <- codes_info %>%
  select(site, location, landuse) %>%
  distinct()


# list of pitfall traps that were sampled 
# compare with this list --> if given sample from this list not present in dataframe of a species group, it means that there were 0 individuals of this group in this sample vs. certain samples got lost and should not be attributed value 0
sampled <- read.csv("Data/pitfall_sampled.csv")
```

The data is processed here, prepared for the analyses.
First, we complete the data with zeroes for organism groups that were sampled in a given trap in a given session, but were not present in the sample.

```{r data}
# read in raw dataframe of harvestmen

harvestmen_raw <- read.csv("Data/Harvestmen.csv") %>%
  select(!c(Stage, Sex)) %>%
  as.data.frame() %>%
  # change from all caps to normal
  mutate(species = str_to_title(species))

# make base dataframe, with species count per session per trap
# complete dataframe and give code to be put as rowname later on
harvestmen <- harvestmen_raw %>%
    # sum "abundances" (activity-density) of same species in same trap
    group_by(site, session, trap, species) %>% # later add site info (location and landuse) again
    reframe(abundance = sum(abundance)) %>%
    ungroup() %>% 
    # right_join with list of sampled "samples" to obtain NA values for study sites that were sampled but did not contain any individuals of this species group
    right_join(sampled %>% select(session, site, trap)) %>%
    # complete with species, so that there is abundance 0 for each species without individuals in this sample (important for species-abundance matrix later on) (would also happen automatically when using pivot_wider)
    complete(nesting(session, site, trap), species) %>%
    # remove "NA" species, no longer necessary (were the result of adding the "empty samples")
    filter(!is.na(species)) %>%
    # abundance NA to 0
    mutate(abundance = replace_na(abundance, 0)) %>%
    # convert session and trap to factor
    mutate(session = factor(session),
           trap = factor(trap)) 

# save
saveRDS(harvestmen, "harvestmen.rds")
```

Now we make three dataframes.
For all the actual analyses, we combine the four sampling sessions, we do not regard them separately.
Because not all sites were visited in all sessions, we average rather than sum the species counts per session.
We can further either a) keep the data per trap, so the average species count per session for each trap; b) combine two traps per site by first summing the species counts per trap and then averaging per session; c) combine two traps per site by taking the average species count per trap and then averaging per session.

If we want to sum two traps per site, it's important that it doesn't occur often that one of the traps is missing for a site.
This is the case in 10 instances out of 169, 1/44 in session 1, 3/44 in session 2, 1/44 in session 3 and 5/44 in session 4. Session 4 does contain the highest proportion of individuals generally (calculated for those traps for which all four sessions were available), so the absence of one of both traps in session 4 can have some influence. On the other hand, this is kind of easier to follow, this approach, clearer to interpret for the reader, and it does not occur thát often.

We will make all three dataframes and perform all the stuff below for all three (for the P2 paper we used data per trap, not combined per site). 

I believe I will choose for option b, but let's see if there are great differences between the three.

```{r threedf}

# per trap
harvestmen_trap <- harvestmen %>%
  # average count per session
  group_by(site, trap, species) %>%
  summarise(count = mean(abundance)) %>%
  # give code
  mutate(code = paste0(site, trap)) %>%
  ungroup() %>%
  select(!c(site, trap))
### checked for AN Oligolophus_tridens and is correctly averaged across sessions per trap
### 6.25 for trap 1 (total count 25) and 0.25 for trap 2 (total count 1)
      
# per site - summed
harvestmen_sitesum <- harvestmen %>%
  # sum two traps per site per session
  group_by(session, site, species) %>%
  reframe(count = sum(abundance)) %>%
  # average count per session
  group_by(site, species) %>%
  summarise(count = mean(count)) %>%
  # temporarily rename site as code (to facilitate loops)
  rename("code" = site)
### checked for AN Oligolophus_tridens and is correctly averaged across sessions per site
### 6.5 (total count 26)
  
# per site - averaged
harvestmen_sitemean <- harvestmen %>%
  # average two traps per site per session
  group_by(session, site, species) %>%
  reframe(count = mean(abundance)) %>%
  # average count per session
  group_by(site, species) %>%
  summarise(count = mean(count)) %>%
  # temporarily rename site as code (to facilitate loops)
  rename("code" = site)
### checked for AN Oligolophus_tridens and is correctly averaged across sessions per site
### 3.25 (6.25 + 0.25)/2


# put the three df in list
dflist <- list("df_trap" = harvestmen_trap,
               "df_sitesum" = harvestmen_sitesum,
               "df_sitemean" = harvestmen_sitemean)

# add one for df_sitesum that doesn't include grasslands and croplands
dflist[["df_sitesum_FOFF"]] <- dflist[["df_sitesum"]] %>%
  left_join(codes_info) %>%
  filter(landuse %in% c("Food_forest", "Forest")) %>%
  select(!c(location, landuse, site)) 

# save
saveRDS(dflist, "dflist.rds")
```

## Data description

```{r data_description}
# # look at distribution per session
# harvestmen %>%
#   group_by(session, site, landuse, trap) %>%
#   # sum counts of different species
#   reframe(total_count = sum(abundance)) %>%
# ggplot(aes(x = session, y = total_count, group = landuse, colour = landuse)) +
#   geom_boxplot() +
#   geom_jitter()
# 
# # look at which proportion, per trap for which all sessions were available, is made up of each session
# testy <- harvestmen %>%
#   group_by(session, site, landuse, trap) %>%
#   # sum counts of different species
#   reframe(total_count = sum(abundance)) %>%
#   # keep only those groups for which the number of sessions equals 4
#   filter(n_distinct(session) == 4) %>%
#   # calculate proportion of each of the sessions (within groups, i.e. traps)
#   ungroup %>%
#   group_by(site, landuse, trap) %>%
#   mutate(prop = 100 * total_count/sum(total_count)) %>%
#   # summary statistics or visual summary of proportion (didn't need this last for visulaisation, e.g. in stacked bar)
#   ungroup() %>%
#   group_by(landuse, session) %>%
#   reframe(mean_prop = mean(prop, na.rm = TRUE))
 
```

## Quantity

Here, for each of the three dataframes, make a dataframe with the ultimate total quantity value per trap or site.
```{r quantitylist}
quantlist <- list()

for(m in names(dflist)){
  quantlist[[m]] <- dflist[[m]] %>%
    ungroup() %>%
    group_by(code) %>% # "trap" not included but can be found as part of "code" for pitfall trap data, and for site rename code to site
    summarise(count = sum(count)) %>% # if we sum first per session and then average sessions or average first and then sum wouldn't make a difference
    # add trap and site info
    left_join(codes_info) %>%
    # landuse to factor
    mutate(landuse = factor(landuse, levels = c("Food_forest",
                                                "Forest",
                                                "Grassland",
                                                "Arable_field")))
}

saveRDS(quantlist, "quantitylist.rds")
```

## Species-abundance matrices

Here, the harvestmen count data is transformed into species-abundance matrices.

For the analysis of the diversity and community composition, juveniles that were unidentified at the species level were excluded, except for Leiobunum_spec. because of these we were sure that they were an extra species not yet present in the samples.
```{r specab}
specablist <- list()

for(m in names(dflist)) {
  specablist[[m]] <- dflist[[m]] %>% 
    # filter out all spec. entries except for Leiobunum_spec.
    filter(!(species %in% c("Phalangiidae_spec.",
                            "Nemastomatidae_spec.", 
                            "Nemastoma_spec."))) %>%
    # make species-abundance matrix
    pivot_wider(names_from = species,
                values_from = count) %>%
  # remove empty communities (necessary for NMDS and PERMANOVA) 
  # for now we don't yet remove all communities that have less than 5 individuals, we just keep all non-empty communities in
    column_to_rownames(var = "code") %>%
    # keep only those rows where the total sum of individuals is > 0 (could be set to different threshold later on)
    filter(rowSums(across(where(is.numeric))) > 0) 
}


saveRDS(specablist, "specablist.rds")
```

## Diversity

As this is based on the specablists, the _spec. entries are already removed.
```{r diversitylist}
diversitylist <- list()

for(m in names(specablist)){
  df <- data.frame(D0 = rowSums(specablist[[m]]!= 0))
  diversitylist[[m]] <- df %>%
  # if 0D is zero, 1D and 2D also need to be zero (if not coded like this, exp(H') = 1 if SR = 0)
    mutate(D1 = ifelse(D0 > 0,
                       hill_taxa(comm = specablist[[m]], q = 1),
                       0),
           D2 = ifelse(D0 > 0,
                       hill_taxa(comm = specablist[[m]], q = 2),
                       0)) %>%
    rownames_to_column(var = "code") %>%
    # re-add information on location and landuse
    left_join(codes_info) %>%
    mutate(landuse = factor(landuse, levels = c("Food_forest",
                                                "Forest",
                                                "Grassland",
                                                "Arable_field")))
}

saveRDS(diversitylist, "diversitylist.rds")
```


## Community composition

### Dissimilarity matrices

Here we produce the Bray-Curtis dissimilarity matrices.

```{r dissimlist}
dissimlist <- list() 

for(g in names(specablist)){
  dissimlist[[g]] <- vegdist(specablist[[g]], method = "bray")
}

saveRDS(dissimlist, "dissimlist.rds")
```


## Compute average pairwise dissimilarity per study site (for beta-diversity)

Not done for this A2 paper.
