---
title: "Prep_communitycomposition"
author: "Lieke"
date: "2024-12-08"
output: html_document
---

```{r packages}
# general
library(tidyverse)
# dissimilarity matrix, NMDS, PERMANOVA, PERMDISP
library(vegan)
```

```{r functions}
source("functions.R")
```

# Data

```{r additional_data}
codes_info <- read.csv("Data/codes_info.csv")
sites_info <- codes_info %>%
  select(site, location, landuse) %>%
  distinct()

grouporder <- c("bacteria", "fungi", "protists", 
                "nematodes", 
                "springtails", "mites", "microarthropods",
                "isopods", "millipedes","centipedes", 
                "harvestmen", "carabids", "macroarthropods",
                "earthworms")

macrofauna_specieslist <- read.csv("Data/macrofauna_specieslist.csv")
```

## Non-sequencing data  

The non-sequencing species data, which is the data of the macrofauna here, is processed here.
First, we complete the data with zeroes for organism groups that were sampled in a given trap in a given session, but were not present in the sample.
```{r data_nonseq}
# read in raw dataframes of macroarthropods
isopods <- read.csv("Data/Isopods_P2.csv") %>%
  as.data.frame() %>%
  # change from all caps to normal
  mutate(species = str_to_title(species))

centipedes <- read.csv("Data/Centipedes_P2.csv") %>%
  as.data.frame()

millipedes <- read.csv("Data/Millipedes_P2.csv") %>%
  as.data.frame()

harvestmen <- read.csv("Data/Harvestmen_P2.csv") %>%
  select(!c(Stage, Sex)) %>%
  as.data.frame() %>%
  # change from all caps to normal
  mutate(species = str_to_title(species))

carabids <- read.csv("Data/Carabids_P2.csv") %>%
  as.data.frame() 

# put dataframes in list
dflist <- list("isopods" = isopods,
               "centipedes" = centipedes,
               "millipedes" = millipedes,
               "harvestmen" = harvestmen,
               "carabids" = carabids)

# read in list of pitfall traps that were sampled 
# compare with this list --> if given sample from this list not present in dataframe of a species group, it means that there were 0 individuals of this group in this sample vs. certain samples got lost and should not be attributed value 0
sampled <- read.csv("Data/pitfall_sampled.csv")

# complete dataframes and give code to be put as rowname later on
for(i in 1:length(dflist)) {
  dflist[[i]] <- dflist[[i]] %>%
    # sum "abundances" (activity-density) of same species in same trap
    group_by(site, session, location, landuse, trap, species) %>% 
    reframe(abundance = sum(abundance)) %>%
    ungroup() %>% 
    # right_join with list of sampled "samples" to obtain NA values for study sites that were sampled but did not contain any individuals of this species group
    right_join(sampled) %>%
    # complete with species, so that there is abundance 0 for each species without individuals in this sample (important for species-abundance matrix later on) (would also happen automatically when using pivot_wider)
    complete(nesting(session, site, location, landuse, trap), species) %>%
    # remove "NA" species, no longer necessary (were the result of adding the "empty samples")
    filter(!is.na(species)) %>%
    # abundance NA to 0
    mutate(abundance = replace_na(abundance, 0)) %>%
    # give code to be used as rowname later on
    mutate(code = paste0(session, site, trap))
}
    
earthworms <- read.csv("Data/Earthworms_P2.csv") %>%
  select(!biomass) %>%
  # for diversity, we don't take juveniles and 'heads' into account (for abundance we do)
  # by doing so, we go from 143 study sites to 136, 7 samples did not contain adults
  filter(!(species %in% c("head", "broken", "juvenile_nopigment", "juvenile_pigment"))) %>%
  # no completion with empty samples needed here, dataframe already contains rows with "empty" for samples that were sampled but did not contain any earthworms
  # complete with species, so that there is abundance 0 for each species without individuals in this sample (important for species-abundance matrix later on)
  complete(nesting(code, site, location, landuse), species) %>%
  replace(is.na(.), 0) %>%
  # remove "empty" rows, not longer needed now that dataframe is completed with 0's for all species
  filter(!(species == "empty"))

# add earthworms to dflist
dflist <- append(dflist, 
                 list("earthworms" = as.data.frame(earthworms)))

# in all dataframes: convert landuse, session and trap to factor
for(g in names(dflist)) {
  dflist[[g]] <- dflist[[g]] %>%
    mutate(landuse = factor(landuse, levels = c("Food_forest",
                                                "Forest",
                                                "Grassland",
                                                "Arable_field"))) %>%
    # also make a factor out of session and trap if these columns exist in the dataframe
    {if("session" %in% colnames(dflist[[g]])) mutate(., session = factor(session)) else .} %>%
    {if("trap" %in% colnames(dflist[[g]])) mutate(., trap = factor(trap)) else .}
}

```

Here, the macrofauna data is transformed into species-abundance matrices.

```{r specab_nonseq}
specablist <- list()

# for isopods, millipedes, centipedes and carabids: averaging species counts of the three first sessions per trap; for harvestmen: averaging species counts of all four sessions per trap
# also already create additional species-abundance matrix for all macroarthropods pooled together
specablist[["macroarthropods"]] <- data.frame(code = paste0(sampled$site, sampled$trap)) %>% distinct()

for(g in c("isopods", "centipedes", "millipedes", "carabids", "harvestmen")) {
  specablist[[g]] <- dflist[[g]] %>% 
    # exclude session 4 in case not harvestmen
    {if(g != "harvestmen") filter(., session != 4) else .} %>%
    # make species-abundance matrix
    pivot_wider(names_from = species,
                values_from = abundance) %>%
    # average across sessions per trap
    group_by(site, trap) %>%
    summarise(across(where(is.numeric), mean)) %>%
    ungroup() %>%
    # give code and remove extra columns
    mutate(code = paste0(site, trap)) %>%
    select(!c(site, trap)) 
  
  specablist[["macroarthropods"]] <- left_join(specablist[["macroarthropods"]],
                                               specablist[[g]], 
                                               by = "code")
  
  # for community composition analyses (PERMANOVA, NMDS, beta-diversity),
  # we remove empty communities, and also all communities that have less than 5 individuals
  specablist[[g]] <- specablist[[g]] %>%
    column_to_rownames(var = "code") %>%
    # keep only those rows where the total sum of individuals is >= 5
    filter(rowSums(across(where(is.numeric))) >= (5/3)) # /3 because we averaged across 3 sessions

  
}

# for community composition analyses (PERMANOVA, NMDS, beta-diversity), 
# we remove empty communities, and also all communities that have less than 5 individuals
specablist[["macroarthropods"]] <- specablist[["macroarthropods"]] %>%
  column_to_rownames(var = "code") %>%
  # keep only those rows where the total sum of individuals is >= 5
  filter(rowSums(across(where(is.numeric))) >= (5/3)) 
  
# for earthworms
specablist[["earthworms"]] <- dflist[["earthworms"]] %>%
      select(c(code, species, abundance)) %>%
      pivot_wider(names_from = species,
                values_from = abundance) %>%
      column_to_rownames(var = "code") %>%
      # keep only those rows where the total sum of individuals is >= 5
      filter(rowSums(across(where(is.numeric))) >= 5)  


# code for averaging two traps (pitfall) and three plots (earthworms) per site can be found in Misc/DA_micro,meso,macrofauna.Rmd
```
Here, we additionally pool the two traps or three plots per site, to be used for the stacked bars that show the average proportion of different taxonomic groups at the site level. We additionally provide extra taxonomic information (genus, family, order, class).

```{r taxon_nonseq}

# macroarthropods

taxon_macroarthropods  <- dflist[["isopods"]] %>%
  mutate(group = "isopods") %>%
  rbind(dflist[["centipedes"]] %>% mutate(group = "centipedes")) %>%
  rbind(dflist[["millipedes"]] %>% mutate(group = "millipedes")) %>%
  rbind(dflist[["carabids"]] %>% mutate(group = "carabids")) %>%
  # for all of these groups filter out session 4
  filter(session != 4) %>%
  # add harvestmen data, for which we use all four sessions
  rbind(dflist[["harvestmen"]] %>% mutate(group = "harvestmen")) %>%
  # calculate average number of individuals per site, averaged across traps and sessions
  group_by(site, group, species) %>%
  reframe(mean_quant = mean(abundance)) %>%
  # add taxonomic information
  left_join(macrofauna_specieslist)

saveRDS(taxon_macroarthropods, "taxon_macroarthropods.rds")

# earthworms

taxon_earthworms <- dflist[["earthworms"]] %>%
  # only keep those communities that consist of at least five individuals (as used for NMDS and PERMANOVA and beta-diversity)
  group_by(code) %>%
  mutate(total_plot_count = sum(abundance)) %>%
  filter(total_plot_count >= 5) %>%
  # calculate mean quantity of each species at the site level
  group_by(site, species) %>%
  reframe(mean_quant = mean(abundance)) %>%
  left_join(macrofauna_specieslist)

saveRDS(taxon_earthworms, "taxon_earthworms.rds")
```

## Dissimilarity matrices

Here we read in the Bray-Curtis dissimilarity matrix for the sequencing data computed on the server (values are the average from 100 dissimilarities, each computed after a random subsample of given read depth was taken from the libraries of the two plots being compared) and produce the Bray-Curtis dissimilarity matrix for the non-sequencing data.

```{r dm_list}
dmlist <- list() # list of dissimilarity matrices for different species groups
```

```{r dissim_matrix_seq}
dmlist[["bacteria"]] <- readRDS("Data/B16S_dist.rds")
dmlist[["fungi"]] <- readRDS("Data/ITS_dist.rds")
dmlist[["protists"]] <- readRDS("Data/P18S_dist_800.rds")
dmlist[["nematodes"]] <- readRDS("Data/N18S_dist.rds")
dmlist[["springtails"]] <- readRDS("Data/COI_col_dist.rds")
# for mites, immediately filter out outlier VB139 here
matrix_mites <- as.matrix(readRDS("Data/COI_mite_dist.rds"))
# filter out outlier
dmlist[["mites"]] <- as.dist(matrix_mites[!rownames(matrix_mites) == "VB139",
                                           !colnames(matrix_mites) == "VB139"])
dmlist[["microarthropods"]] <- readRDS("Data/COI_ma_dist_1000.rds")
```

```{r dissim_matrix_nonseq}
for(g in names(specablist)){
  dmlist[[g]] <- vegdist(specablist[[g]], method = "bray")
}

dmlist <- dmlist[grouporder]

saveRDS(dmlist, "dmlist.rds")
```

## Compute average pairwise dissimilarity per study site (for beta-diversity)

```{r avgdissim_within}
avgdissimlist <- list()
# per species group
for(g in names(dmlist)){
  df <- dmlist[[g]]
  avgdissimlist[[g]] <- data.frame()
   # the below creates a dataframe that links codes (that are rownames in the dissimilarity matrix) to sites, so that we can do grouping per site
  rn <- data.frame(code = rownames(as.matrix(df))) %>%
    left_join(codes_info %>% select(code, site))

  avg_pairwise_dm <- meandist(df, grouping = rn$site) # diagonal is within-site mean dissim

  output <- data.frame(site = names(diag(avg_pairwise_dm)),
                      avg_pairwise_dissim = diag(avg_pairwise_dm)) %>%
    left_join(codes_info %>% select(site, location, landuse) %>% distinct()) %>%
    # remove NA-rows
    filter(!is.na(avg_pairwise_dissim))

  avgdissimlist[[g]] <- rbind(avgdissimlist[[g]], output)
}

saveRDS(avgdissimlist, "avgdissimlist.rds")
```

### Is it an issue for average pairwise dissimilarity that the number of sample plots/traps per study site is not always the same?
For each species group the number of plots/traps that was sampled for this species group is the same in every study site, BUT: some plots/traps did not contain individuals of this species group, so that the number of plots/traps based on which the average pairwise dissimilarity is calculated is not the same
- for sequencing data: presumably every plot contains ASVs for each species group, so OK
- for non-sequencing data: 
a) pitfall trap data: only two traps, so if one was always empty (all three sessions), the dissimilarity cannot be calculated anyway, so this wouldn't pose a problem for this (i.e. both if one or if two of the traps have total abundance of 0, this study site is left out for this dissimilarity analysis)
b) earthworms: here, there are a few cases where only 1/3 plots did not contain at least 5 individuals, which results in its average pairwise dissimilarity being calculated based on 2 rather than 3 plots (so just equal to the BC dissimilarity between those two plots). This is the case for: ... (should check this again now that we use minimum of 5 individuals)
So for Malden and Vaassen forests, all plots were empty, for Sint-Truiden 2/3 plots, and these study sites are therefore left out of this analysis. But for ..., only 1/3 plots "contained" less than 5 individuals so that the avg pairwise dissimilarity is computed based on 2 plots only. I would not consider this a problem.

We should check if the patterns in overall dissimilarity show a relation with patterns in total abundance/...
